---
title: "Tracking the mental health of Australian workforce 2024-25"
author: Dev Gur
format: html
editor: visual
code-fold: true
---

## Introduction

Workplace mental health represents a complex interaction between individual psychological symptoms and the broader organisational environment in which employees operate. As Australian workers face evolving pressures in the post-pandemic landscape, understanding how psychological distress, burnout, and organisational climate factors fluctuate over time has become critical for developing effective intervention strategies. This analysis examines quarterly trends across 2024 and 2025 to determine whether psychological distress (K6), burnout (BAT-12), SuperFriend domains, and psychosocial hazards have shifted systematically over this two-year period. Particular attention is given to the Retail and Transport industries to assess whether sector-specific change is evident, as these industries face distinct operational pressures and workforce compositions. To ensure robust inference, three complementary approaches are triangulated: all possible quarter-to-quarter comparisons with Bonferroni correction to control Type I error, linear regression models assessing gradual temporal trends, and chi-square tests evaluating changes in the proportion of workers meeting clinical distress thresholds. This multi-method approach strengthens confidence in observed patterns and provides a comprehensive evidence base for organisational and policy responses.

```{r message=FALSE}
library(tidyverse)
library(haven)
library(forcats)
library(effsize)  
library(broom)
library(dplyr)

# Load SPSS files
itw_2024 <- read_sav("data/ITW_2024.sav") |> zap_labels()
itw_2025 <- read_sav("data/ITW_2025.sav") |> zap_labels()

# Harmonise variable names
itw_2025 <- itw_2025 |> rename(K6score = k6score)

# Combine datasets
itw_all <- bind_rows(itw_2024, itw_2025) |>
  mutate(
    Year = as.integer(as.character(Year)),
    Year = ifelse(is.na(Year), 2025L, Year),
    Wave = as.integer(as.character(Wave)),

    Quarter = factor(
      paste0("Q", Wave),
      levels = c("Q1","Q2","Q3","Q4"),
      ordered = TRUE
    ),

    Quarter_num = as.numeric(Quarter)
  )

# Save clean dataset
saveRDS(itw_all, "data/itw_all_clean.rds")


```

```{r include=FALSE}
# NATIONAL DATASET 
itw_national <- itw_all |>
  mutate(Group = "National")

# INDUSTRY DATASET (Retail and Transport)
itw_industry <- itw_all |>
  mutate(
    Group = case_when(
      a6 == 7 ~ "Retail",
      a6 == 9 ~ "Transport",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(Group))

itw_industry |> count(Group)
```

### Superfriend domain scores

```{r}
connectedness_items <- paste0("b1_", 1:10)
leadership_items    <- paste0("b2_", 1:10)

safety_items <- c(
  "b3x1_1",
  paste0("b3x2_", 1:3),
  paste0("b3x3_", 1:3),
  paste0("b3x4_", 1:3)
)

work_design_items <- c(
  paste0("b4x1_", 1:3),
  paste0("b4x2_", 1:4),
  paste0("b4x3_", 1:3)
)

capability_items <- paste0("b5_", 1:7)

score_sf <- function(df, items) {

  m <- rowMeans(df[, items], na.rm = TRUE)
  s <- (m - 1) * 25
  s[is.nan(m)] <- NA
  s
}


itw_national <- itw_national |>
  mutate(
    Connectedness = score_sf(itw_national, connectedness_items),
    Leadership    = score_sf(itw_national, leadership_items),
    Safety        = score_sf(itw_national, safety_items),
    Work_Design   = score_sf(itw_national, work_design_items),
    Capability    = score_sf(itw_national, capability_items)
  )

itw_industry <- itw_industry |>
  mutate(
    Connectedness = score_sf(itw_industry, connectedness_items),
    Leadership    = score_sf(itw_industry, leadership_items),
    Safety        = score_sf(itw_industry, safety_items),
    Work_Design   = score_sf(itw_industry, work_design_items),
    Capability    = score_sf(itw_industry, capability_items)
  )


# Hazard items
mgmt_support_items <- c("b2_1","b2_2","b2_3","b2_4","b2_5")
workload_items     <- c("b4x1_1","b4x1_2","b4x1_3")
recognition_items  <- c("b2_6","b2_7","b2_8","b2_9","b2_10")

# Function to calculate hazard scores
score_hazard <- function(df, items){
  m <- rowMeans(df[, items], na.rm=TRUE)
  s <- m * 20
  s[is.nan(m)] <- NA
  return(s)
}

# Score hazards for industry
itw_industry <- itw_industry |>
  mutate(
    Hazard_Management_Support = score_hazard(itw_industry, mgmt_support_items),
    Hazard_Workload           = score_hazard(itw_industry, workload_items),
    Hazard_Recognition        = score_hazard(itw_industry, recognition_items)
  )

```

### Psychological hazards

```{r}
mgmt_support_items <- c("b2_1","b2_2","b2_3","b2_4","b2_5")
workload_items     <- c("b4x1_1","b4x1_2","b4x1_3")
recognition_items  <- c("b2_6","b2_7","b2_8","b2_9","b2_10")



itw_national <- itw_national |>
  mutate(
    Hazard_Management_Support = score_hazard(itw_national, mgmt_support_items),
    Hazard_Workload           = score_hazard(itw_national, workload_items),
    Hazard_Recognition        = score_hazard(itw_national, recognition_items)
  )
```

### Aggregation helper

```{r}
agg_qtr_year <- function(df, var){
  df |>
    group_by(Year, Quarter) |>
    summarise(
      mean = mean(.data[[var]], na.rm=TRUE),
      sd   = sd(.data[[var]], na.rm=TRUE),
      n    = sum(!is.na(.data[[var]])),
      se   = sd/sqrt(n),
      lci  = mean - 1.96*se,
      uci  = mean + 1.96*se,
      .groups="drop"
    )
}
```

```{r}
# Convert all outcome columns to numeric
itw_national <- itw_national |> 
  mutate(across(
    c(K6score, BAT12_total,
      Connectedness, Leadership, Safety, Work_Design, Capability,
      Hazard_Management_Support, Hazard_Workload, Hazard_Recognition),
    ~ as.numeric(.)
  ))

itw_industry <- itw_industry |>
  mutate(across(
    c(K6score, BAT12_total,
      Connectedness, Leadership, Safety, Work_Design, Capability,
      Hazard_Management_Support, Hazard_Workload, Hazard_Recognition),
    ~ as.numeric(.)
  ))
```

```{r}
numeric_vars <- c(
  "K6score","BAT12_total",
  "Connectedness","Leadership","Safety","Work_Design","Capability",
  "Hazard_Management_Support","Hazard_Workload","Hazard_Recognition"
)

itw_national <- itw_national |> mutate(across(all_of(numeric_vars), as.numeric))
itw_industry <- itw_industry |> mutate(across(all_of(numeric_vars), as.numeric))
```

## National Plots

**K6**

```{r}
k6_nat <- agg_qtr_year(itw_national, "K6score")

ggplot(k6_nat, aes(x = Quarter, y = mean,
                   colour = factor(Year),
                   group = Year)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = lci, ymax = uci), width = 0.1) +
  scale_colour_manual(values = c("blue","red")) +
  theme_minimal() +
  labs(
    title = "Psychological Distress (K6) – National Workforce",
    x = "Quarter",
    y = "Mean K6 (95% CI)",
    colour = "Year",
    caption = "Note. Higher K6 scores indicate greater psychological distress. Error bars represent 95% confidence intervals."
  )


```

2025 shows marginally lower distress levels across the year compared to 2024, indicating some improvement, distress never returns to the exceptionally low Q1 baseline, remaining elevated throughout the remainder of both years. The confidence intervals tighten in later quarters, suggesting that as distress rises, worker experiences become more consistent rather than more variable.

**Is there any seasonal affect for psychological distress for 2024 and 2025?** There is evidence suggestive of a seasonal pattern. In both 2024 and 2025, psychological distress peaks in Q2, with lower levels observed in Q1. This recurring mid year elevation indicates a potential seasonal effect. However, patterns diverge in the second half of the year. While distress declines gradually after Q2 in 2024, 2025 shows a sharper drop in Q3 followed by a slight increase in Q4. Because the late year trajectories differ between years, the data indicate a recurring Q2 peak in psychological distress, but not a perfectly stable seasonal cycle across all quarters.

**Burnout**

```{r}
bat_nat <- agg_qtr_year(itw_national, "BAT12_total")

ggplot(bat_nat, aes(x = Quarter, y = mean,
                    colour = factor(Year),
                    group = Year)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = lci, ymax = uci), width = 0.1) +
  scale_colour_manual(values = c("blue","red")) +
  theme_minimal() +
  labs(
    title = "Burnout (BAT-12) – National Workforce",
    x = "Quarter",
    y = "Mean BAT-12 (95% CI)",
    colour = "Year",
    caption = "Note. Higher BAT-12 scores indicate greater burnout. Error bars represent 95% confidence intervals."
  )


```

Burnout levels in the national workforce show a relatively stable pattern across quarters in both 2024 and 2025. Mean BAT-12 scores vary within a narrow range across the calendar year, indicating only minimal changes in overall burnout levels. In both years, burnout increases slightly from Q1 to Q2. In 2024, levels continue to rise in Q3 before declining slightly in Q4, while in 2025 burnout decreases in Q3 and then increases again in Q4. Although these patterns show some variation across the second half of the year, the overall differences between quarters remain small. Because this figure represents the overall BAT-12 score, it reflects general burnout levels and does not distinguish between individual components of burnout. Overall, the results suggest that national burnout levels remained broadly consistent across the study period, with only minimal quarter to quarter variation.

**Superfriend domains- National**

```{r}
sf_domains <- c("Connectedness","Leadership","Safety",
                "Work_Design","Capability")

sf_nat <- itw_national |>
  select(Year, Quarter, all_of(sf_domains)) |>
  pivot_longer(
    -c(Year, Quarter),
    names_to = "Domain",
    values_to = "Score"
  ) |>
  mutate(
    Domain = factor(
      Domain,
      levels = c("Capability", "Work_Design", "Leadership", "Safety", "Connectedness")
    )
  ) |>
  group_by(Year, Quarter, Domain) |>
  summarise(
    mean = mean(Score, na.rm = TRUE),
    se   = sd(Score, na.rm = TRUE) / sqrt(sum(!is.na(Score))),
    lci  = mean - 1.96 * se,
    uci  = mean + 1.96 * se,
    .groups = "drop"
  )

ggplot(
  sf_nat,
  aes(Quarter, mean, colour = factor(Year), group = Year)
) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = lci, ymax = uci), width = .1) +
  facet_wrap(~Domain, ncol = 3) +
  scale_colour_manual(values = c("blue", "red")) +
  theme_minimal() +
  labs(
    title = "SuperFriend Domains – National",
    y = "Mean (95% CI)",
    colour = "Year"
  )

```

The five Domains of a Thriving Workplace present a mixed picture, Leadership shows a slightly downward trajectory across quarters, while Safety, Work Design, Connectedness, and Capability remain relatively consistent across both years. Overall, the results suggest that the Domains were largely stable over time, with no substantial shifts in domain scores across quarters.

**Psychosocial Hazards**

```{r}
haz_vars <- c(
  "Hazard_Management_Support",
  "Hazard_Workload",
  "Hazard_Recognition"
)

haz_nat <- itw_national |>
  select(Year, Quarter, all_of(haz_vars)) |>
  pivot_longer(-c(Year,Quarter),
               names_to="Hazard",
               values_to="Score") |>
  group_by(Year,Quarter,Hazard) |>
  summarise(
    mean=mean(Score,na.rm=TRUE),
    se=sd(Score,na.rm=TRUE)/sqrt(sum(!is.na(Score))),
    lci=mean-1.96*se,
    uci=mean+1.96*se,
    .groups="drop"
  )

ggplot(haz_nat,
       aes(Quarter,mean,
           colour=factor(Year),
           group=Year))+
  geom_line()+
  geom_point()+
  geom_errorbar(aes(ymin=lci,ymax=uci),width=.1)+
  facet_wrap(~Hazard)+
  theme_minimal()+
  labs(
    title="Psychosocial Hazards – National",
    y="Mean Score",
    colour="Year"
  )

```

Psychosocial hazard indicators show similar patterns across quarters for Management Support and Recognition. In both for both of these hazards, scores increase from Q1 to Q2 and then gradually decline through to Q4, indicating a consistent mid year peak followed by a steady decrease. This pattern is observed in both 2024 and 2025, suggesting a comparable trend across years. While some sequential quarter to quarter comparisons show statistically significant differences, these changes are relatively modest in size and should be interpreted cautiously. In contrast, Workload remains relatively stable across quarters in both years, with only small fluctuations over time. Overall, the results indicate that psychosocial hazard scores show moderate and consistent patterns across quarters, with some modest gradual changes and relatively little variation in Workload across the national workforce.

## Industry Plots

**K6**

```{r}
k6_ind <- itw_industry |>
  group_by(Group, Year, Quarter) |>
  summarise(
    mean = mean(K6score, na.rm=TRUE),
    se   = sd(K6score, na.rm=TRUE)/sqrt(sum(!is.na(K6score))),
    lci  = mean - 1.96*se,
    uci  = mean + 1.96*se,
    .groups="drop"
  )

k6_national_avg <- mean(itw_national$K6score, na.rm = TRUE)

ggplot(k6_ind,
       aes(Quarter, mean,
           colour = factor(Year),
           group = Year)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=lci,ymax=uci),width=.1) +
  geom_hline(
  yintercept = k6_national_avg,
  linetype = "dashed",
  colour = "black"
) +
  facet_wrap(~Group) +
  scale_colour_manual(values = c("blue","red")) +
  theme_minimal() +
  labs(
    title="K6 – Retail vs Transport",
    x = "Quarter",
  y = "Mean K6 (95% CI)",
  colour = "Year",
  caption = "Note. Higher K6 scores indicate higher levels of psychological distress. Error bars represent 95% CI."
)


```

The Retail industry shows greater variation in psychological distress across quarters compared with Transport. In Retail, distress increases from Q1 to Q2 in both years, with Q2 representing the highest mean levels of distress. Distress then declines in Q3 and Q4 of 2025, while remaining relatively elevated in 2024. Notably, mean K6 scores in Retail appear higher in Q3 and Q4 of 2024 compared with the same quarters in 2025. In contrast, Transport shows relatively stable distress levels across quarters in both years, with smaller changes between adjacent quarters.

**Burnout**

```{r}
# Burnout – Industry

bat_ind <- itw_industry |>
  group_by(Group, Year, Quarter) |>
  summarise(
    mean = mean(BAT12_total, na.rm=TRUE),
    se   = sd(BAT12_total, na.rm=TRUE)/sqrt(sum(!is.na(BAT12_total))),
    lci  = mean - 1.96*se,
    uci  = mean + 1.96*se,
    .groups="drop"
  )

ggplot(bat_ind,
       aes(Quarter, mean,
           colour = factor(Year),
           group = Year)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=lci,ymax=uci),width=.1) +
  facet_wrap(~Group) +
  scale_colour_manual(values=c("blue","red")) +
  theme_minimal() +
  labs(
    title="Burnout (BAT-12) – Retail vs Transport",
    x = "Quarter",
    y="Mean BAT-12 (95% CI)",
    colour="Year",
     caption = "Note. Higher BAT-12 scores indicate higher levels of psychological distress. Error bars represent 95% CI."
  )

```

Retail shows moderate variation in burnout scores across quarters in both years. In Retail, burnout increases from Q1 to Q2 in both 2024 and 2025, followed by a decline in Q3 before rising slightly again in Q4. In several quarters, particularly Q3, mean burnout levels appear higher in 2024 than in 2025. Transport shows a comparatively stable pattern, with mean BAT-12 scores remaining within a relatively narrow range across quarters in both years. Confidence intervals overlap substantially across quarters and between years, suggesting that many of the observed differences are relatively small. Overall, the figure indicates modest fluctuations in burnout across quarters, with somewhat greater variability in Retail compared with Transport.

```{r}
itw_national <- itw_national |>
  mutate(K6_modsev = ifelse(K6score>=13,1,0))

itw_industry <- itw_industry |>
  mutate(K6_modsev = ifelse(K6score>=13,1,0))

```

### Chi square test industry

```{r include=FALSE}
chi_results <- itw_industry |>
  group_by(Group) |>
  summarise(

    chisq_p =
      chisq.test(table(Quarter,K6_modsev))$p.value,

    trend_p = {

      tbl <- table(Quarter,K6_modsev)

      prop.trend.test(tbl[,2],
                      rowSums(tbl))$p.value
    },

    .groups="drop"
  )

chi_results

```

**Retail:** There was a statistically significant change in moderate to severe psychological distress across quarters in Retail χ² p \< 0.001, with a significant upward linear trend p = 0.0003.

**Transport:** No significant change in psychological distress across quarters was observed in Transport (p = 0.901).

### NATIONAL t-test

```{r include=FALSE}
itw_national <- itw_national |> 
  mutate(Quarter = factor(Quarter, levels=c("Q1","Q2","Q3","Q4"), ordered=TRUE))


# Outcomes
outcomes <- c(
  "K6score","BAT12_total",
  "Connectedness","Leadership","Safety",
  "Work_Design","Capability",
  "Hazard_Management_Support",
  "Hazard_Workload","Hazard_Recognition"
)

# Function: Pairwise tests + manual mean difference
run_pairwise_ttests <- function(df, outcome) {
  
  df <- df |>
    select(Quarter, all_of(outcome)) |>
    drop_na()
  
  # Compute quarter means
  means_tbl <- df |>
    group_by(Quarter) |>
    summarise(mean_value = mean(.data[[outcome]]),
              .groups = "drop")
  
  # Pairwise p-values
  pw <- pairwise.t.test(
    x = df[[outcome]],
    g = df$Quarter,
    p.adjust.method = "none"
  ) |>
    broom::tidy()
  
  # Add mean difference manually
  pw <- pw |>
    rowwise() |>
    mutate(
      estimate =
        means_tbl$mean_value[means_tbl$Quarter == group2] -
        means_tbl$mean_value[means_tbl$Quarter == group1]
    ) |>
    ungroup() |>
    mutate(outcome = outcome)
  
  return(pw)
}

# Run tests separately for each year
ttest_nat_all <- bind_rows(
  lapply(c(2024, 2025), function(yr){
    
    bind_rows(
      lapply(outcomes, function(out){
        
        run_pairwise_ttests(
          filter(itw_national, Year == yr),
          out
        ) |>
          mutate(Year = yr)
        
      }))
    }))

# Bonferroni correction (within Year + Outcome)
ttest_nat_all <- ttest_nat_all |>
  group_by(Year, outcome) |>
  mutate(
    p_adj = p.adjust(p.value, method = "bonferroni"),
    significant = p_adj < 0.05
  ) |>
  ungroup()

ttest_nat_all
```

### Mental Health Outcomes (K6 + BAT12)

```{r}
mh_vars <- c("K6score","BAT12_total")

ggplot(
  ttest_nat_all |> 
    filter(significant == TRUE,
           outcome %in% mh_vars),
  aes(x = paste(group1, "vs", group2),
      y = estimate,
      fill = outcome)
) +
  geom_col(position = "dodge") +
  facet_wrap(~Year) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Significant Quarter-wise Comparisons – Mental Health",
    x = "Quarter Comparison",
    y = "Mean Difference (Group2 - Group1)",
    fill = "Outcome"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The bar chart reveals that most significant differences in mental health outcomes occur when comparing later quarters to Q1, particularly in 2024. Psychological distress K6 shows consistent positive mean differences when comparing Q2, Q3, and Q4 against Q1, indicating distress accumulates and peaks as the year progresses. Burnout (BAT12) follows a similar pattern but with smaller magnitude changes. Notably, 2025 shows fewer significant quarter to quarter differences, suggesting a potential stabilisation or plateau in mental health deterioration during the second year. The consistent pattern of positive differences confirms that both distress and burnout are significantly higher in later quarters compared to Q1, supporting the presence of a cumulative effect across the working year.

### SuperFriend Domains

```{r}
domain_vars <- c("Connectedness","Leadership",
                 "Safety","Work_Design","Capability")

ggplot(
  ttest_nat_all |> 
    filter(significant == TRUE,
           outcome %in% domain_vars),
  aes(x = paste(group1, "vs", group2),
      y = estimate,
      fill = outcome)
) +
  geom_col(position = "dodge") +
  facet_wrap(~Year) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Significant Quarter-wise Comparisons – SuperFriend Domains",
    x = "Quarter Comparison",
    y = "Mean Difference (Group2 - Group1)",
    fill = "Domain"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Leadership emerges as the most dynamic domain, showing multiple significant quarter-to-quarter fluctuations. The largest positive difference appears in Q4 vs Q2 comparisons for Leadership, suggesting some late-year improvement in leadership perceptions, though this follows earlier declines. Connectedness shows smaller but significant variations, primarily in mid year comparisons. Work Design improvements appear predominantly in 2025, indicating potential structural workplace adjustments taking effect in the second year. The absence of Safety from significant comparisons is noteworthy - this domain remained remarkably stable across all quarters, suggesting safety perceptions may be more resistant to short term fluctuations or that safety protocols are consistently maintained regardless of other organisational changes.

### Psychosocial Hazards

```{r}
haz_vars <- c("Hazard_Management_Support",
              "Hazard_Workload",
              "Hazard_Recognition")

ggplot(
  ttest_nat_all |> 
    filter(significant == TRUE,
           outcome %in% haz_vars),
  aes(x = paste(group1, "vs", group2),
      y = estimate,
      fill = outcome)
) +
  geom_col(position = "dodge") +
  facet_wrap(~Year, scales = "free_y") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Significant Quarter-wise Comparisons – Psychosocial Hazards",
    x = "Quarter Comparison",
    y = "Mean Difference (Group2 - Group1)",
    fill = "Hazard"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Recognition and Management Support show the most volatile patterns, with significant negative differences in mid year comparisons Q2 vs Q1, Q3 vs Q2 indicating deterioration in these critical supervisory domains during the first three quarters. The y-axis scales across facets reveal that Hazard_Workload shows virtually no significant quarter to quarter variation, reinforcing the regression finding that workload remained stable while relational factors deteriorated. Interestingly, 2025 shows fewer significant fluctuations in hazards, which may indicate either adaptation or resignation to sustained poor conditions. The pattern suggests that organisational culture factors recognition, support are more seasonally variable than operational demands workload, which remain relatively constant throughout the year.

### Industry t-test

```{r include=FALSE, warning=FALSE}
itw_industry <- itw_industry |>
  mutate(
    Quarter = factor(Quarter,
                     levels = c("Q1","Q2","Q3","Q4"),
                     ordered = TRUE),
    Quarter_num = as.numeric(Quarter)
  )

ttest_industry_all <- bind_rows(
  lapply(c("Retail","Transport"), function(ind){
    
    lapply(c(2024, 2025), function(yr){
      
      bind_rows(
        lapply(outcomes, function(out){
          
          run_pairwise_ttests(
            filter(itw_industry,
                   Group == ind,
                   Year == yr),
            out
          ) |>
            mutate(Group = ind,
                   Year = yr)
          
        })
      )
      
    }) |> bind_rows()
    
  })
)

ttest_industry_all <- ttest_industry_all |>
  group_by(Group, Year, outcome) |>
  mutate(
    p_adj = p.adjust(p.value, method = "bonferroni"),
    significant = p_adj < 0.05
  ) |>
  ungroup()

ttest_industry_all
```

```{r warning=FALSE}

ggplot(
  ttest_industry_all |> filter(significant == TRUE),
  aes(x = paste(group1, "vs", group2),
      y = estimate,
      fill = outcome)
) +
  geom_col() +
  facet_grid(Group ~ Year, scales = "free_y") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Significant Quarter-wise Pairwise Comparisons – Industry",
    x = "Quarter Comparison",
    y = "Mean Difference (Group2 - Group1)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )

```

Retail shows substantially more significant quarter to quarter variations than Transport, with Burnout dominating the significant comparisons. Retail demonstrates significant differences, confirming the cumulative deterioration pattern identified in the regression analysis. Safety appears as a significant domain only in Retail, suggesting this industry experiences meaningful fluctuations in perceived psychological safety that Transport workers do not. The concentration of significant findings in Retail confirms this sector's particular vulnerability to temporal fluctuations in workforce mental health.

### Time series regression National

```{r include=FALSE}
run_lm_trend <- function(df,outcome){

  lm(as.formula(paste(outcome,"~ Quarter_num")),
     data=df) |>
    tidy() |>
    filter(term=="Quarter_num") |>
    mutate(outcome=outcome)

}

mh_nat <- bind_rows(

  run_lm_trend(itw_national,"K6score"),

  run_lm_trend(itw_national,"BAT12_total")

)

mh_nat

```

**K6:** Psychological distress increased significantly over time nationally β = 0.099 per quarter, p = 0.001.

**Burnout:** Burnout scores also increased significantly across quarters β = 0.013, p = 0.002.

### Domain scores regression

```{r include=FALSE}
sf_nat_reg <- bind_rows(

  run_lm_trend(itw_national,"Connectedness"),
  run_lm_trend(itw_national,"Leadership"),
  run_lm_trend(itw_national,"Safety"),
  run_lm_trend(itw_national,"Work_Design"),
  run_lm_trend(itw_national,"Capability")

)

sf_nat_reg

```

Only Leadership showed a statistically meaningful change. Linear regression analysis showed a significant decline in Leadership domain scores over time β = −0.259 per quarter, p = 0.018, indicating worsening perceptions of leadership support across 2024–2025. No statistically significant time trends were observed for Connectedness, Safety, or Work Design since p \> 0.05. Capability showed a non-significant upward trend p = 0.093.

### Psychosocial hazards regression

```{r include=FALSE}
haz_nat_reg <- bind_rows(

  run_lm_trend(itw_national,"Hazard_Management_Support"),
  run_lm_trend(itw_national,"Hazard_Workload"),
  run_lm_trend(itw_national,"Hazard_Recognition")

)

haz_nat_reg

```

Management Support scores declined significantly over time β = −0.193, p = 0.033, as did Recognition scores β = −0.0111, p = 0.014, suggesting worsening psychosocial working conditions nationally. No significant change was observed for Workload p = 0.709.

### Industry regression

```{r include=FALSE}
mh_industry <- bind_rows(

  run_lm_trend(
    filter(itw_industry,Group=="Retail"),
    "K6score"
  ) |> mutate(Group="Retail"),

  run_lm_trend(
    filter(itw_industry,Group=="Transport"),
    "K6score"
  ) |> mutate(Group="Transport")

)

mh_industry

```

**Retail:** Psychological distress increased significantly over time in Retail β = 0.358, p = 0.001.

**Transport:** No significant time trend was observed in Transport p = 0.298.

## Conclusion

Across 2024–2025, this analysis reveals a consistent and concerning pattern of deteriorating workforce mental health in Australia, driven primarily by relational and organisational factors rather than task demands. Psychological distress increased significantly nationally, with a pronounced seasonal peak in Q2 of both years from Q1 baseline that suggests predictable mid year crisis points aligned with end of financial year pressures. Burnout followed a similar upward trend but showed remarkable stability across quarters, indicating chronic exhaustion that changes slowly despite acute distress fluctuations. Critically, while workload remained constant, leadership perceptions declined significantly, accompanied by deteriorating management support and recognition. This dissociation between stable task demands and worsening relational factors confirms that the deterioration stems from how workers are treated, not how much work they perform. Industry comparisons revealed stark disparities- Retail workers experienced significant distress increases and multiple significant quarter to quarter fluctuations, while Transport workers remained stable throughout, suggesting sector specific vulnerabilities related to customer facing stress, understaffing, and precarious employment arrangements. The convergence of pairwise testing, regression modelling, and categorical trend analysis provides consistent evidence that workforce mental health deterioration is both real and targeted worsening overall but concentrated in specific sectors and driven by manageable organisational factors. These findings point clearly toward interventions focused on leadership development, recognition programs, and Q2 specific support, particularly within the Retail sector, rather than simplistic workload reduction strategies.
