---
title: "Analysis of Psychological and Financial distress in the Australian Workforce"
author: Dev Gur
format: html
editor: visual
code-fold: true
---

## Introduction

This analysis examines how financial distress among Australian workers relates to broader economic conditions by integrating external macroeconomic indicators with survey data. In particular, the study incorporates the Australian Bureau of Statistics (ABS) Living Cost Index (LCI) and the Reserve Bank of Australia (RBA) cash rate to explore whether changes in financial wellbeing align with shifts in cost of living and monetary policy over time.

The ABS Living Cost Index, sourced from the official ABS publication Living Cost Indexes for Selected Household Types, measures changes in the cost of maintaining a constant standard of living for different household groups. This analysis uses the Employee households LCI, as it most closely reflects the financial pressures experienced by working Australians. The index captures variations in essential expenses such as housing, utilities, and food, providing a more targeted measure of household cost pressures compared to broader inflation indicators.

In addition, the RBA cash rate, obtained from the Reserve Bank of Australia’s official data releases, represents a key monetary policy instrument that influences borrowing costs, mortgage repayments, and overall household financial conditions. Changes in interest rates can therefore have direct implications for financial stress, particularly among working age individuals with debt or housing commitments.

Both datasets were aligned with the survey data on a quarterly basis for the period 2024–2025, enabling direct comparison with financial distress measures derived from the ITW dataset. By combining these sources, the analysis aims to assess whether financial distress trends correspond with changes in living costs and interest rates, and to identify how these relationships vary across different demographic and occupational groups.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(haven)
library(janitor)
library(readxl)
library(lubridate)
library(ggplot2)

# Read and clean ITW files
itw_2024 <- read_sav("data/ITW_2024.sav") |>
  clean_names()

itw_2025 <- read_sav("data/ITW_2025.sav") |>
  clean_names()

# Harmonise variable names
itw_2024 <- itw_2024 |>
  rename(k6_score = k6score)

itw_2025 <- itw_2025 |>
  rename(k6_score = k6score)

# Add year
itw_2024 <- itw_2024 |>
  mutate(year = 2024)

itw_2025 <- itw_2025 |>
  mutate(year = 2025)

# Combine
itw_all <- bind_rows(itw_2024, itw_2025)

# Fix missing year values and create quarter
itw_all <- itw_all |>
  mutate(
    year = if_else(is.na(year) & wave == 4, 2025, year),
    quarter = paste(year, paste0("Q", wave)),
    age_group = as_factor(age)
  )

# Check counts
itw_all |>
  count(year, wave)

# Create national quarterly summary
wave_summary <- itw_all |>
  group_by(quarter) |>
  summarise(
    mean_f10 = mean(f10_1, na.rm = TRUE),
    high_f10_pct = mean(f10_1 >= 8, na.rm = TRUE) * 100,
    mean_k6 = mean(k6_score, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

# Read ABS file
abs_raw <- read_excel(
  "data/allgroupsbyhouseholds.xlsx",
  sheet = "Data1",
  col_names = FALSE
) |>
  clean_names()

# Data rows start at row 11
date_row_idx <- 11

# Extract Employee households all-groups index
abs_lci <- abs_raw |>
  slice(date_row_idx:n()) |>
  transmute(
    date_raw = as.numeric(x1),
    employee_lci = as.numeric(x3)
  ) |>
  mutate(
    date = as.Date(date_raw, origin = "1899-12-30"),
    quarter = paste(format(date, "%Y"), paste0("Q", quarter(date)))
  ) |>
  select(quarter, employee_lci)

# Keep 2024-2025 only
abs_lci_2425 <- abs_lci |>
  filter(quarter %in% c(
    "2024 Q1", "2024 Q2", "2024 Q3", "2024 Q4",
    "2025 Q1", "2025 Q2", "2025 Q3", "2025 Q4"
  ))

# Merge
phase2_dataset <- wave_summary |>
  left_join(abs_lci_2425, by = "quarter")

# Result
phase2_dataset
colSums(is.na(phase2_dataset))
```

### Mapping the cost of living data with financial distress

```{r warning=FALSE}


ggplot(phase2_dataset, aes(x = quarter)) +
  geom_line(aes(y = mean_f10, group = 1), color = "blue", size = 1.2) +
  geom_point(aes(y = mean_f10), color = "blue", size = 3) +
  geom_line(aes(y = employee_lci/20, group = 2), color = "red", size = 1.2) +
  geom_point(aes(y = employee_lci/20), color = "red", size = 3) +
  labs(
    title = "Financial Distress and Living Cost Index Over Time",
    x = "Quarter",
    y = "Financial Distress (scaled)",
    caption = "ABS Selected Living Cost Index – Employee households"
  ) +
  theme_minimal()
```

The national trend highlights a clear and meaningful relationship between financial distress and living costs over time. While the Living Cost Index shows a steady upward movement, financial distress exhibits moderate fluctuations, suggesting that workers are actively adjusting to ongoing economic pressures. The alignment between the two series indicates that cost of living changes are an important contextual factor influencing workforce financial wellbeing.

```{r}
cor_test_national <- cor.test(
  phase2_dataset$mean_f10,
  phase2_dataset$employee_lci
)

cor_test_national
```

The correlation between financial distress and the living cost index is negative and not statistically significant (p \> 0.05). This suggests that there is no strong linear relationship at the national level over this time period.

```{r}
age_summary <- itw_all |>
  filter(!is.na(age_group)) |>
  group_by(quarter, age_group) |>
  summarise(
    mean_f10 = mean(f10_1, na.rm = TRUE),
    n = sum(!is.na(f10_1)),
    .groups = "drop"
  )

age_summary
```

```{r}
age_phase2 <- age_summary |>
  left_join(abs_lci_2425, by = "quarter")

age_phase2
```

```{r}
age_phase2 <- age_phase2 |>
  group_by(age_group) |>
  mutate(
    f10_z = scale(mean_f10)[,1],
    lci_z = scale(employee_lci)[,1]
  ) |>
  ungroup()
```

```{r}
ggplot(age_phase2, aes(x = quarter, group = age_group)) +
  # Real 1-10 Distress Score
  geom_line(aes(y = mean_f10, colour = "Financial Distress"), linewidth = 1.1) +
  geom_point(aes(y = mean_f10, colour = "Financial Distress"), size = 2) +
  
  # LCI divided by 20 (The "Giant to Ant" fix)
  geom_line(aes(y = employee_lci/20, colour = "Living Cost Index (scaled)"), 
            linewidth = 1.1, linetype = "dashed") +
  
  facet_wrap(~age_group) +
  scale_colour_manual(values = c("Financial Distress" = "blue", 
                                 "Living Cost Index (scaled)" = "red")) +
  labs(
    title = "Financial Distress and Living Cost Index by Age Group",
    subtitle = "LCI scaled by factor of 20 for visual comparison",
    x = "Quarter",
    y = "Score / Scaled Index",
    colour = "",
    caption = "Higher F10 scores indicate greater financial distress."
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

The age based analysis provides valuable insight into how different demographic groups experience financial pressure. Younger and working age groups (18–44 years) consistently show higher levels of financial distress and greater variation over time, reflecting their stronger exposure to housing costs, debt and economic changes. In contrast, older age groups display more stable and lower distress levels, indicating greater financial resilience. This highlights the importance of targeted support for younger workers.

```{r}
age_cor <- age_phase2 |>
  group_by(age_group) |>
  summarise(
    correlation = cor(mean_f10, employee_lci, use = "complete.obs"),
    .groups = "drop"
  )

age_cor
```


## High Financial stress category
```{r}
financial_stress_summary <- itw_all |>
  filter(!is.na(financial_stress)) |>
  group_by(quarter) |>
  summarise(
    high_stress_pct = mean(financial_stress == 2, na.rm = TRUE) * 100,
    low_stress_pct  = mean(financial_stress == 1, na.rm = TRUE) * 100,
    n = n(),
    .groups = "drop"
  )

financial_stress_summary
```

```{r}
ggplot(financial_stress_summary, aes(x = quarter)) +
  geom_line(aes(y = high_stress_pct, group = 1), color = "red", linewidth = 1.2) +
  geom_point(aes(y = high_stress_pct), color = "red", size = 3) +
  labs(
    title = "Percentage of Workers with High Financial Stress Over Time",
    x = "Quarter",
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```

The trend in high financial stress provides a focused view of workforce vulnerability. Peaks in early 2024 highlight periods of heightened pressure, followed by a gradual decline into 2025. This suggests that workers are adapting over time and that financial stress is responsive to broader economic conditions, demonstrating improving stability in later periods.


## Psychological vs Financial Distress

```{r}
#| label: lag-calc
phase2_lagged <- phase2_dataset %>%
  mutate(lci_lag1 = lag(employee_lci))

# Correlation with 1-quarter lag
cor_lag <- cor.test(phase2_lagged$mean_f10, phase2_lagged$lci_lag1, use = "complete.obs")
```

```{r warning=FALSE}
ggplot(phase2_lagged, aes(x = lci_lag1, y = mean_f10)) +
  geom_point(color = "purple", size = 3) +
  geom_smooth(method = "lm", color = "black") +
  labs(
    title = "Time-Lagged Effect: Costs vs Distress",
    subtitle = "Does last quarter's cost predict this quarter's stress?",
    x = "Living Cost Index (Previous Quarter)", 
    y = "Mean Financial Distress (Current Quarter)"
  ) +
  theme_minimal()
```
It visually confirms that what   correlation test were hinting at, it's a negative relationship. The downward sloping line suggests that as the Living Cost Index (LCI) from the previous quarter increases, the Mean Financial Distress in the current quarter actually appears to trend downward.
There could be a few possibilities here-

- Distress often peaks when a crisis is new (uncertainty), rather than when it is at its highest point (reality).

- Many Australian workers received pay rises or adjusted their tax brackets in late 2024/early 2025, which may have "buffered" the rising costs.



```{r}
#| label: psych-vs-fin
ggplot(phase2_dataset, aes(x = mean_f10, y = mean_k6)) +
  geom_point(aes(size = n), color = "darkgreen", alpha = 0.6) +
  geom_smooth(method = "lm", color = "black", linetype = "dashed") +
  labs(
    title = "Correlation: Financial vs Psychological Distress",
    subtitle = "National Average 2024-2025",
    x = "Mean Financial Distress (F10)",
    y = "Mean Psychological Distress (K6)",
    caption = "Higher F10 and K6 scores indicate greater distress. Line represents linear regression"
  ) +
  theme_minimal()

# Calculate Pearson correlation
cor_psych <- cor(phase2_dataset$mean_f10, phase2_dataset$mean_k6)
```

The relationship between financial and psychological distress shows a clear and consistent positive association. Higher financial distress aligns with increased psychological distress, reinforcing the close connection between economic conditions and mental wellbeing. This highlights the importance of addressing financial challenges as part of broader workforce mental health strategies.

## Occupational Differences- Blue collar vs White collar

```{r}
# RQ#5: Blue Collar (2) vs White Collar (1)
occupational_summary <- itw_all |>
  filter(!is.na(a18), a18 %in% c(1, 2)) |>
  mutate(worker_type = if_else(a18 == 2, "Blue Collar", "White Collar")) |>
  group_by(quarter, worker_type) |>
  summarise(
    mean_f10 = mean(f10_1, na.rm = TRUE),
    mean_k6 = mean(k6_score, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  left_join(abs_lci_2425, by = "quarter")
```

```{r}
worker_correlations <- occupational_summary |>
  group_by(worker_type) |>
  summarise(
    f10_lci_cor = cor(mean_f10, employee_lci, use = "complete.obs"),
    .groups = "drop"
  )

print(worker_correlations)
```

White collar has strong negative relation which means that as cost of living rise, their distress actually fell. For Blue collar, it has much weaker correlation, their distress doesn't drop when there is shift in the economy.

```{r}
ggplot(occupational_summary, aes(x = quarter, group = worker_type)) +
  # Real 1-10 Distress Score
  geom_line(aes(y = mean_f10, color = "Financial Distress"), linewidth = 1.2) +
  geom_point(aes(y = mean_f10), color = "blue", size = 2) +
  
  # LCI divided by 20
  geom_line(aes(y = employee_lci/20, color = "Living Cost Index (scaled)"), 
            linewidth = 1.2, linetype = "dashed") +
  
  facet_wrap(~worker_type) +
  scale_color_manual(values = c("Financial Distress" = "blue", 
                                "Living Cost Index (scaled)" = "red")) +
  labs(
    title = "Living Costs vs Financial Distress: Blue vs White Collar",
    subtitle = "Distress (1-10) vs. LCI (scaled by /20)",
    x = "Quarter",
    y = "Financial distress"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The comparison across occupational groups reveals distinct and insightful patterns. Blue collar workers experience slightly higher and more variable financial distress, indicating greater sensitivity to economic changes. White collar workers show more stable trends, suggesting stronger buffering against cost pressures. These differences highlight how job type influences financial wellbeing and the importance of tailored workplace support

## Cost of Living Pressures (Housing, Food, Utilities)

```{r}
f8_summary <- itw_all |>
  select(quarter, f8_1, f8_2, f8_3) |>
  pivot_longer(cols = starts_with("f8"), 
               names_to = "category", 
               values_to = "rating") |>
  filter(!is.na(rating)) |>
  mutate(
    category = case_when(
      category == "f8_1" ~ "Housing (Rent/Mortgage)",
      category == "f8_2" ~ "Utilities(Electricity/Gas/Water)",
      category == "f8_3" ~ "Food"
    ),
    # Convert numeric 1-5 to readable labels
    rating_label = factor(rating, 
                          levels = 1:5, 
                          labels = c("1 - Very Difficult", "2 - Difficult", "3 - Neutral", "4 - Easy", "5 - Very Easy"))
  ) |>
  group_by(quarter, category, rating_label) |>
  summarise(count = n(), .groups = "drop") |>
  group_by(quarter, category) |>
  mutate(pct = (count / sum(count)) * 100)
```

```{r}
ggplot(f8_summary, aes(x = quarter, y = pct, fill = rating_label)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~category) +
  scale_fill_brewer(palette = "RdYlGn") + # Red (Difficult) to Green (Easy)
  labs(
    title = "Ease of Meeting Essential Living Costs",
    subtitle = "Comparing Housing, Utilities, and Food across 2024-2025",
    x = "Quarter",
    y = "Percentage of Workers (%)",
    fill = "Ease of Payment"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  scale_y_continuous(labels = scales::percent_format())
```

The breakdown of essential living costs provides a comprehensive view of everyday financial experiences. Across all categories, a substantial proportion of workers report moderate to high ease of payment, reflecting a degree of financial adaptability. However, housing consistently shows somewhat greater difficulty compared to food and utilities, reinforcing its role as the primary financial pressure point for households.

## RBA Interest rates Comparison

```{r}
# Create RBA Cash Rate data for 2024-2025
rba_rates <- tibble(
  quarter = c("2024 Q1", "2024 Q2", "2024 Q3", "2024 Q4", 
              "2025 Q1", "2025 Q2", "2025 Q3", "2025 Q4"),
  cash_rate = c(4.35, 4.35, 4.35, 4.35, 4.10, 3.85, 3.60, 3.60)
)

# Merge with your existing phase 2 data
phase2_with_rba <- phase2_dataset |>
  left_join(rba_rates, by = "quarter")
```

```{r}
ggplot(phase2_with_rba, aes(x = quarter)) +
  # 1. Financial Distress (Raw 1-10)
  geom_line(aes(y = mean_f10, group = 1, color = "Financial Distress"), linewidth = 1.2) +
  geom_point(aes(y = mean_f10), color = "blue", size = 3) +
  
  # 2. RBA Cash Rate 
  
  geom_line(aes(y = cash_rate, group = 2, color = "RBA Cash Rate (%)"), 
            linewidth = 1.2, linetype = "dashed") +
  geom_point(aes(y = cash_rate), color = "orange", size = 3) +
  
  scale_color_manual(values = c("Financial Distress" = "blue", "RBA Cash Rate (%)" = "orange")) +
  
  labs(
    title = "Financial Distress vs RBA Interest Rates",
    subtitle = "Comparing Workforce Sentiment with Monetary Policy",
    x = "Quarter",
    y = "Score",
    color = "Indicator"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```

The comparison between financial distress and RBA interest rates reveals as interest rates begin to ease in 2025, financial distress levels also stabilise and slightly decline. This parallel movement suggests that monetary policy plays a meaningful role in shaping financial wellbeing, with improvements in economic conditions reflected in workforce sentiment.

```{r}
# Calculate correlation between Financial Distress and RBA Cash Rate
cor_rba <- cor.test(
  phase2_with_rba$mean_f10, 
  phase2_with_rba$cash_rate, 
  use = "complete.obs"
)

# Print the result
cor_rba
```
The analysis reveals a strong positive directional trend. While the cost of living (ABS) showed a complex, adaptive relationship with workers, the RBA interest rates shows a much more direct impact. We found a strong 0.62 correlation, suggesting that the cost of debt is perhaps a more immediate driver of workforce anxiety than the general rise in grocery or utility prices.

```{r}
library(dplyr)
# Merge age summary with RBA rates
age_rba_analysis <- age_summary |>
  left_join(rba_rates, by = "quarter")

# Quick check to ensure cash_rate is present
head(age_rba_analysis)
```
```{r}
ggplot(age_rba_analysis, aes(x = quarter, group = age_group)) +
  # 1. Financial Distress by Age (Blue Line)
  geom_line(aes(y = mean_f10, color = "Financial Distress"), linewidth = 1.1) +
  geom_point(aes(y = mean_f10, color = "Financial Distress"), size = 2) +
  
  # 2. RBA Cash Rate Benchmark (Orange Dashed Line)
  geom_line(aes(y = cash_rate, color = "RBA Cash Rate (%)"), 
            linewidth = 1, linetype = "dashed") +
  
  facet_wrap(~age_group) +
  
  # 3. Colors and Labels
  scale_color_manual(values = c("Financial Distress" = "blue", 
                                 "RBA Cash Rate (%)" = "orange")) +
  labs(
    title = "Age-Based Financial Distress vs. RBA Interest Rates",
    subtitle = "Comparing generational sentiment against the national cost of debt",
    x = "Quarter",
    y = "Score / Percentage",
    color = "Indicator",
    caption = "Note. F10 (Distress) and RBA (%) are raw scores. Data: ITW & RBA (2024-2025)."
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    legend.position = "bottom",
    strip.text = element_text(face = "bold") # Makes age group titles pop
  )
```

```{r}
# 1. Create the Org Size summary
org_summary <- itw_all |>
  filter(!is.na(a5)) |>
  mutate(org_size = as_factor(a5)) |> # Ensure a5 is a readable factor
  group_by(quarter, org_size) |>
  summarise(mean_f10 = mean(f10_1, na.rm = TRUE), .groups = "drop")

# 2. Join with RBA rates
org_rba_analysis <- org_summary |>
  left_join(rba_rates, by = "quarter")

# 3. Plot it
ggplot(org_rba_analysis, aes(x = quarter, group = org_size)) +
  geom_line(aes(y = mean_f10, color = "Financial Distress"), linewidth = 1.1) +
  geom_line(aes(y = cash_rate, color = "RBA Cash Rate (%)"), linetype = "dashed") +
  facet_wrap(~org_size) +
  scale_color_manual(values = c("Financial Distress" = "blue", "RBA Cash Rate (%)" = "orange")) +
  theme_minimal() +
  labs(title = "Financial Distress vs. RBA Rates by Org Size",
       caption = "Note. F10 and RBA are raw scores. Data: ITW & RBA (2024-25).") +
   theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
   )
```
### National Trend Significance: A Descriptive Decline

```{r}
ggplot(phase2_dataset, aes(x = as.numeric(as.factor(quarter)), y = mean_f10)) +
  geom_point(size = 3, color = "blue") +
  geom_smooth(method = "lm", color = "darkblue", fill = "lightblue") +
  labs(
    title = "National Financial Distress Trend Significance",
    subtitle = "Linear regression over 8 quarters (2024-2025)",
    x = "Time Period (Quarters)",
    y = "Mean Financial distress Score",
    caption = "Note. Shaded area represents the 95% confidence interval."
  ) +
  theme_minimal()
```

There is a negative linear trend, which means national financial distress has generally decreased from its peak in early 2024.

The 95% confidence interval (the light blue shaded area) is relatively wide, which indicates some quarterly volatility. However, the consistent downward slope suggest that the Australian workforce is transitioning from a period of some shock to one of gradual adaptation as we move into 2026.

### Macroeconomic Pressure: Interest Rates vs Essential Utilities
```{r}
# 1. Summarize Utilities (f8_2)
utilities_summary <- itw_all |>
  group_by(quarter) |>
  summarise(util_difficulty = mean(f8_2, na.rm = TRUE)) |>
  left_join(rba_rates, by = "quarter")

# 2. Plot it
ggplot(utilities_summary, aes(x = cash_rate, y = util_difficulty)) +
  geom_point(size = 4, color = "orange") +
  geom_smooth(method = "lm", color = "black", linetype = "dashed") +
  labs(
    title = "Monetary Policy vs. Utility Pressure",
    x = "RBA Cash Rate (%)",
    y = "Ease of Paying Utilities (1-5 scale)"
  ) +
  theme_minimal()

# 3. Correlation
cor.test(utilities_summary$cash_rate, utilities_summary$util_difficulty)
```

There is a clear negative correlation ($R^2 = 0.49$, representing a large effect size) between interest rates and the ease of paying for electricity, gas, and water. As the RBA Cash Rate increases, workers report significantly more difficulty meeting their utility costs.

This statistically supports the finding that high interest rates don't just affect mortgages,  they squeeze the household budget so tightly that even basic essentials like utilities become harder to manage. This is a key insight for SuperFriend’s advocacy work."


### Linear Regression model
```{r}
# 1. Join RBA rates to your Occupational Summary
model_data <- occupational_summary |>
  left_join(rba_rates, by = "quarter") |>
  # Create the "Dummy" variable: 0 = White Collar, 1 = Blue Collar
  mutate(D = if_else(worker_type == "Blue Collar", 1, 0))

# 2. Run the Linear Model
# Formula: Financial Distress ~ RBA Rate + Dummy + (RBA Rate * Dummy)
dummy_model <- lm(mean_f10 ~ cash_rate * D, data = model_data)

# 3. See the results
summary(dummy_model)
```

When interest rates go up, everyone feels the squeeze. While Blue Collar workers start with higher financial stress overall, they don't necessarily get hit harder by rate hikes than White Collar workers, both groups financial stress levels move up and down at the same pace.

Our model explained 56% of why people's financial stress levels change, which is a very high score. It shows that interest rates are a universal pressure that affects the whole Australian workforce. This is a powerful tool for SuperFriend because now we can actually predict when workers will need more support based on what the RBA does with interest rates.

Basically, the economy is the volume knob for workforce financial stress, when the RBA turns it up, the pressure rises for everyone across the board.




## Conclusion

Overall, the analysis provides a clear and insightful view of how financial distress among Australian workers evolves in relation to broader economic conditions. At the national level, financial distress shows moderate variation over time, while the Living Cost Index demonstrates a steady upward trend. Although the direct statistical relationship between these variables is not strong, the visual patterns suggest that financial distress responds to changing economic conditions in a dynamic and adaptive manner.

The comparison with RBA interest rates further highlights the influence of macroeconomic factors, with financial distress stabilising as interest rates begin to ease in 2025. This indicates that improvements in economic conditions may be reflected in workforce financial wellbeing over time.

Disaggregated analysis reveals meaningful differences across population groups. Younger and working age individuals, particularly those aged 25–44 years, display greater sensitivity to cost of living changes, reflecting their higher exposure to housing costs and financial commitments. In contrast, older age groups show more stable trends, suggesting greater financial resilience. Similarly, differences across occupational groups indicate that financial experiences are not uniform, with some groups exhibiting greater variability in response to economic pressures.

The analysis of financial stress categories shows that the majority of workers remain in the low stress group, while fluctuations in high financial stress appear to be short term and responsive rather than persistent. Additionally, the strong alignment between financial and psychological distress highlights the close connection between economic conditions and overall wellbeing.

In summary, the findings suggest that while cost of living pressures and interest rate changes influence financial distress, the Australian workforce demonstrates resilience and adaptability over time. The variation observed across age groups, occupations, and stress levels underscores the importance of targeted, data driven approaches to supporting financial and mental wellbeing in different segments of the workforce.

