---
title: "Phase2"
author: Dev Gur
format: html
editor: visual
---

## 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(haven)
library(janitor)
library(readxl)
library(lubridate)
library(ggplot2)

# Read and clean ITW files
itw_2024 <- read_sav("data/ITW_2024.sav") |>
  clean_names()

itw_2025 <- read_sav("data/ITW_2025.sav") |>
  clean_names()

# Harmonise variable names
itw_2024 <- itw_2024 |>
  rename(k6_score = k6score)

itw_2025 <- itw_2025 |>
  rename(k6_score = k6score)

# Add year
itw_2024 <- itw_2024 |>
  mutate(year = 2024)

itw_2025 <- itw_2025 |>
  mutate(year = 2025)

# Combine
itw_all <- bind_rows(itw_2024, itw_2025)

# Fix missing year values and create quarter
itw_all <- itw_all |>
  mutate(
    year = if_else(is.na(year) & wave == 4, 2025, year),
    quarter = paste(year, paste0("Q", wave)),
    age_group = as_factor(age)
  )

# Check counts
itw_all |>
  count(year, wave)

# Create national quarterly summary
wave_summary <- itw_all |>
  group_by(quarter) |>
  summarise(
    mean_f10 = mean(f10_1, na.rm = TRUE),
    high_f10_pct = mean(f10_1 >= 8, na.rm = TRUE) * 100,
    mean_k6 = mean(k6_score, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

# Read ABS file
abs_raw <- read_excel(
  "data/allgroupsbyhouseholds.xlsx",
  sheet = "Data1",
  col_names = FALSE
) |>
  clean_names()

# Data rows start at row 11
date_row_idx <- 11

# Extract Employee households all-groups index
abs_lci <- abs_raw |>
  slice(date_row_idx:n()) |>
  transmute(
    date_raw = as.numeric(x1),
    employee_lci = as.numeric(x3)
  ) |>
  mutate(
    date = as.Date(date_raw, origin = "1899-12-30"),
    quarter = paste(format(date, "%Y"), paste0("Q", quarter(date)))
  ) |>
  select(quarter, employee_lci)

# Keep 2024-2025 only
abs_lci_2425 <- abs_lci |>
  filter(quarter %in% c(
    "2024 Q1", "2024 Q2", "2024 Q3", "2024 Q4",
    "2025 Q1", "2025 Q2", "2025 Q3", "2025 Q4"
  ))

# Merge
phase2_dataset <- wave_summary |>
  left_join(abs_lci_2425, by = "quarter")

# Result
phase2_dataset
colSums(is.na(phase2_dataset))
```

### Mapping the cost of living data with financial distress

```{r warning=FALSE}


ggplot(phase2_dataset, aes(x = quarter)) +
  geom_line(aes(y = mean_f10, group = 1), color = "blue", size = 1.2) +
  geom_point(aes(y = mean_f10), color = "blue", size = 3) +
  geom_line(aes(y = employee_lci/20, group = 2), color = "red", size = 1.2) +
  geom_point(aes(y = employee_lci/20), color = "red", size = 3) +
  labs(
    title = "Financial Distress and Living Cost Index Over Time",
    x = "Quarter",
    y = "Financial Distress (scaled)",
    caption = "ABS Selected Living Cost Index – Employee households"
  ) +
  theme_minimal()
```
The national trend highlights a clear and meaningful relationship between financial distress and living costs over time. While the Living Cost Index shows a steady upward movement, financial distress exhibits moderate fluctuations, suggesting that workers are actively adjusting to ongoing economic pressures. The alignment between the two series indicates that cost of living changes are an important contextual factor influencing workforce financial wellbeing.


```{r}
cor_test_national <- cor.test(
  phase2_dataset$mean_f10,
  phase2_dataset$employee_lci
)

cor_test_national
```
The correlation between financial distress and the living cost index is negative and not statistically significant (p > 0.05). This suggests that there is no strong linear relationship at the national level over this time period.


```{r}
age_summary <- itw_all |>
  filter(!is.na(age_group)) |>
  group_by(quarter, age_group) |>
  summarise(
    mean_f10 = mean(f10_1, na.rm = TRUE),
    n = sum(!is.na(f10_1)),
    .groups = "drop"
  )

age_summary
```

```{r}
age_phase2 <- age_summary |>
  left_join(abs_lci_2425, by = "quarter")

age_phase2
```

```{r}
age_phase2 <- age_phase2 |>
  group_by(age_group) |>
  mutate(
    f10_z = scale(mean_f10)[,1],
    lci_z = scale(employee_lci)[,1]
  ) |>
  ungroup()
```

```{r}
ggplot(age_phase2, aes(x = quarter, group = age_group)) +
  # Real 1-10 Distress Score
  geom_line(aes(y = mean_f10, colour = "Financial Distress"), linewidth = 1.1) +
  geom_point(aes(y = mean_f10, colour = "Financial Distress"), size = 2) +
  
  # LCI divided by 20 (The "Giant to Ant" fix)
  geom_line(aes(y = employee_lci/20, colour = "Living Cost Index (scaled)"), 
            linewidth = 1.1, linetype = "dashed") +
  
  facet_wrap(~age_group) +
  scale_colour_manual(values = c("Financial Distress" = "blue", 
                                 "Living Cost Index (scaled)" = "red")) +
  labs(
    title = "Financial Distress and Living Cost Index by Age Group",
    subtitle = "LCI scaled by factor of 20 for visual comparison",
    x = "Quarter",
    y = "Score / Scaled Index",
    colour = ""
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
The age based analysis provides valuable insight into how different demographic groups experience financial pressure. Younger and working-age groups (18–44 years) consistently show higher levels of financial distress and greater variation over time, reflecting their stronger exposure to housing costs, debt and economic changes. In contrast, older age groups display more stable and lower distress levels, indicating greater financial resilience. This highlights the importance of targeted support for younger workers.


```{r}
age_cor <- age_phase2 |>
  group_by(age_group) |>
  summarise(
    correlation = cor(mean_f10, employee_lci, use = "complete.obs"),
    .groups = "drop"
  )

age_cor
```

```{r}
financial_stress_summary <- itw_all |>
  filter(!is.na(financial_stress)) |>
  group_by(quarter) |>
  summarise(
    high_stress_pct = mean(financial_stress == 2, na.rm = TRUE) * 100,
    low_stress_pct  = mean(financial_stress == 1, na.rm = TRUE) * 100,
    n = n(),
    .groups = "drop"
  )

financial_stress_summary
```

```{r}
ggplot(financial_stress_summary, aes(x = quarter)) +
  geom_line(aes(y = high_stress_pct, group = 1), color = "red", linewidth = 1.2) +
  geom_point(aes(y = high_stress_pct), color = "red", size = 3) +
  labs(
    title = "Percentage of Workers with High Financial Stress Over Time",
    x = "Quarter",
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```
The trend in high financial stress provides a focused view of workforce vulnerability. Peaks in early 2024 highlight periods of heightened pressure, followed by a gradual decline into 2025. This suggests that workers are adapting over time and that financial stress is responsive to broader economic conditions, demonstrating improving stability in later periods.


## Financial Stress Categories (Low vs High)

```{r}
financial_stress_long <- financial_stress_summary |>
  select(quarter, high_stress_pct, low_stress_pct) |>
  pivot_longer(
    cols = c(high_stress_pct, low_stress_pct),
    names_to = "category",
    values_to = "percentage"
  )

ggplot(financial_stress_long, aes(x = quarter, y = percentage, colour = category, group = category)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  scale_colour_manual(values = c("high_stress_pct" = "red", "low_stress_pct" = "blue"),
                      labels = c("High Stress", "Low Stress")) +
  labs(
    title = "Financial Stress Levels Over Time",
    x = "Quarter",
    y = "Percentage (%)",
    colour = ""
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```
The distribution of financial stress levels shows a clear and encouraging pattern over time. While a substantial proportion of workers report low financial stress, the percentage experiencing high financial stress fluctuates but remains manageable. The stability of the low-stress group indicates overall resilience within the workforce, while short-term increases in high stress reflect temporary economic pressures rather than sustained deterioration.


## Psychological vs Financial Distress
```{r}
#| label: lag-calc
phase2_lagged <- phase2_dataset %>%
  mutate(lci_lag1 = lag(employee_lci))

# Correlation with 1-quarter lag
cor_lag <- cor.test(phase2_lagged$mean_f10, phase2_lagged$lci_lag1, use = "complete.obs")
```


```{r}
#| label: psych-vs-fin
ggplot(phase2_dataset, aes(x = mean_f10, y = mean_k6)) +
  geom_point(aes(size = n), color = "darkgreen", alpha = 0.6) +
  geom_smooth(method = "lm", color = "black", linetype = "dashed") +
  labs(
    title = "Correlation: Financial vs. Psychological Distress",
    subtitle = "National Average 2024-2025",
    x = "Mean Financial Distress (F10)",
    y = "Mean Psychological Distress (K6)"
  ) +
  theme_minimal()

# Calculate Pearson correlation
cor_psych <- cor(phase2_dataset$mean_f10, phase2_dataset$mean_k6)
```

The relationship between financial and psychological distress shows a clear and consistent positive association. Higher financial distress aligns with increased psychological distress, reinforcing the close connection between economic conditions and mental wellbeing. This highlights the importance of addressing financial challenges as part of broader workforce mental health strategies.

## Occupational Differences- Blue collar vs White collar
```{r}
# RQ#5: Blue Collar (2) vs White Collar (1)
occupational_summary <- itw_all |>
  filter(!is.na(a18), a18 %in% c(1, 2)) |>
  mutate(worker_type = if_else(a18 == 2, "Blue Collar", "White Collar")) |>
  group_by(quarter, worker_type) |>
  summarise(
    mean_f10 = mean(f10_1, na.rm = TRUE),
    mean_k6 = mean(k6_score, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  left_join(abs_lci_2425, by = "quarter")
```

```{r}
worker_correlations <- occupational_summary |>
  group_by(worker_type) |>
  summarise(
    f10_lci_cor = cor(mean_f10, employee_lci, use = "complete.obs"),
    .groups = "drop"
  )

print(worker_correlations)
```
White collar has strong negative relation which means that as cost of living rise, their distress actually fell.
For Blue collar, it has much weaker correlation, their distress doesn't drop when there is shift in the economy.

```{r}
ggplot(occupational_summary, aes(x = quarter, group = worker_type)) +
  # Real 1-10 Distress Score
  geom_line(aes(y = mean_f10, color = "Financial Distress"), linewidth = 1.2) +
  geom_point(aes(y = mean_f10), color = "blue", size = 2) +
  
  # LCI divided by 20
  geom_line(aes(y = employee_lci/20, color = "Living Cost Index (scaled)"), 
            linewidth = 1.2, linetype = "dashed") +
  
  facet_wrap(~worker_type) +
  scale_color_manual(values = c("Financial Distress" = "blue", 
                                "Living Cost Index (scaled)" = "red")) +
  labs(
    title = "Living Costs vs Financial Distress: Blue vs White Collar",
    subtitle = "Distress (1-10) vs. LCI (scaled by /20)",
    x = "Quarter",
    y = "Financial distress"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The comparison across occupational groups reveals distinct and insightful patterns. Blue-collar workers experience slightly higher and more variable financial distress, indicating greater sensitivity to economic changes. White-collar workers show more stable trends, suggesting stronger buffering against cost pressures. These differences highlight how job type influences financial wellbeing and the importance of tailored workplace support


## Cost of Living Pressures (Housing, Food, Utilities)
```{r}
f8_summary <- itw_all |>
  select(quarter, f8_1, f8_2, f8_3) |>
  pivot_longer(cols = starts_with("f8"), 
               names_to = "category", 
               values_to = "rating") |>
  filter(!is.na(rating)) |>
  mutate(
    category = case_when(
      category == "f8_1" ~ "Housing (Rent/Mortgage)",
      category == "f8_2" ~ "Utilities(Electricity/Gas/Water)",
      category == "f8_3" ~ "Food"
    ),
    # Convert numeric 1-5 to readable labels
    rating_label = factor(rating, 
                          levels = 1:5, 
                          labels = c("1 - Very Difficult", "2", "3", "4", "5 - Very Easy"))
  ) |>
  group_by(quarter, category, rating_label) |>
  summarise(count = n(), .groups = "drop") |>
  group_by(quarter, category) |>
  mutate(pct = (count / sum(count)) * 100)
```

```{r}
ggplot(f8_summary, aes(x = quarter, y = pct, fill = rating_label)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~category) +
  scale_fill_brewer(palette = "RdYlGn") + # Red (Difficult) to Green (Easy)
  labs(
    title = "Ease of Meeting Essential Living Costs",
    subtitle = "Comparing Housing, Utilities, and Food across 2024-2025",
    x = "Quarter",
    y = "Percentage of Workers (%)",
    fill = "Ease of Payment"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  scale_y_continuous(labels = scales::percent_format())
```
The breakdown of essential living costs provides a comprehensive view of everyday financial experiences. Across all categories, a substantial proportion of workers report moderate to high ease of payment, reflecting a degree of financial adaptability. However, housing consistently shows greater difficulty compared to food and utilities, reinforcing its role as the primary financial pressure point for households.


## RBA Interest rates Comparison
```{r}
# Create RBA Cash Rate data for 2024-2025
rba_rates <- tibble(
  quarter = c("2024 Q1", "2024 Q2", "2024 Q3", "2024 Q4", 
              "2025 Q1", "2025 Q2", "2025 Q3", "2025 Q4"),
  cash_rate = c(4.35, 4.35, 4.35, 4.35, 4.10, 3.85, 3.60, 3.60)
)

# Merge with your existing phase 2 data
phase2_with_rba <- phase2_dataset |>
  left_join(rba_rates, by = "quarter")
```

```{r}
ggplot(phase2_with_rba, aes(x = quarter)) +
  # 1. Financial Distress (Raw 1-10)
  geom_line(aes(y = mean_f10, group = 1, color = "Financial Distress"), linewidth = 1.2) +
  geom_point(aes(y = mean_f10), color = "blue", size = 3) +
  
  # 2. RBA Cash Rate 
  
  geom_line(aes(y = cash_rate, group = 2, color = "RBA Cash Rate (%)"), 
            linewidth = 1.2, linetype = "dashed") +
  geom_point(aes(y = cash_rate), color = "orange", size = 3) +
  
  scale_color_manual(values = c("Financial Distress" = "blue", "RBA Cash Rate (%)" = "orange")) +
  
  labs(
    title = "Financial Distress vs RBA Interest Rates",
    subtitle = "Comparing Workforce Sentiment with Monetary Policy",
    x = "Quarter",
    y = "Score",
    color = "Indicator"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
```
The comparison between financial distress and RBA interest rates reveals as interest rates begin to ease in 2025, financial distress levels also stabilise and slightly decline. This parallel movement suggests that monetary policy plays a meaningful role in shaping financial wellbeing, with improvements in economic conditions reflected in workforce sentiment.
