---
title: "Analysis of Financial and Psychological Distress in the Australian Workforce"
subtitle: "Tracking workforce mental health, financial pressure, living costs and interest rates"
author: "Dev Gur"
date: "25 May 2026"

title-slide-attributes:
  data-background-gradient: "linear-gradient(135deg, #071A2C 0%, #173A59 50%, #00A7C8 100%)"

format:
  revealjs:
    theme: simple
    transition: fade
    slide-number: true
    code-fold: true
    width: 1400
    height: 800
    css: styles.css
    center-title-slide: false

execute:
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(haven)
library(janitor)
library(forcats)
library(scales)
library(readxl)
library(lubridate)
```

```{r data-prep, include=FALSE}
# ------------------------------------------------------------
# 1. Find your ITW 2024 and 2025 .sav files
# ------------------------------------------------------------

all_sav_files <- list.files(
  path = ".",
  pattern = "\\.sav$",
  recursive = TRUE,
  full.names = TRUE,
  ignore.case = TRUE
)

file_2024 <- all_sav_files[str_detect(all_sav_files, "2024")][1]
file_2025 <- all_sav_files[str_detect(all_sav_files, "2025")][1]

if (is.na(file_2024) | is.na(file_2025)) {
  stop("Could not find both 2024 and 2025 .sav files. Put them in this project folder or a data folder.")
}

# ------------------------------------------------------------
# 2. Helper functions
# ------------------------------------------------------------

read_itw <- function(path, year_value) {
  read_sav(path) |>
    clean_names() |>
    mutate(year = year_value)
}

pick_num <- function(df, possible_names) {
  selected_name <- possible_names[possible_names %in% names(df)][1]
  
  if (is.na(selected_name)) {
    return(rep(NA_real_, nrow(df)))
  }
  
  as.numeric(df[[selected_name]])
}

score_mean_100 <- function(df, item_names) {
  item_names <- intersect(item_names, names(df))
  
  if (length(item_names) == 0) {
    return(rep(NA_real_, nrow(df)))
  }
  
  mean_score <- rowMeans(
    as.data.frame(lapply(df[item_names], as.numeric)),
    na.rm = TRUE
  )
  
  score_100 <- (mean_score - 1) * 25
  score_100[is.nan(mean_score)] <- NA_real_
  
  score_100
}

ci_summary <- function(df, variable, group_vars) {
  df |>
    group_by(across(all_of(group_vars))) |>
    summarise(
      mean = mean(.data[[variable]], na.rm = TRUE),
      sd = sd(.data[[variable]], na.rm = TRUE),
      n = sum(!is.na(.data[[variable]])),
      se = sd / sqrt(n),
      lower = mean - 1.96 * se,
      upper = mean + 1.96 * se,
      .groups = "drop"
    )
}

# ------------------------------------------------------------
# 3. Read and combine data
# ------------------------------------------------------------

itw_2024 <- read_itw(file_2024, 2024)
itw_2025 <- read_itw(file_2025, 2025)

itw_all <- bind_rows(itw_2024, itw_2025)

# ------------------------------------------------------------
# 4. Create variables used in slides
# ------------------------------------------------------------

itw_all <- itw_all |>
  mutate(
    wave = pick_num(cur_data(), c("wave", "quarter", "timepoint")),
    quarter = factor(
      paste0("Q", wave),
      levels = c("Q1", "Q2", "Q3", "Q4"),
      ordered = TRUE
    ),
    
    k6_score = pick_num(cur_data(), c("k6score", "k6_score", "k6_total")),
    bat12_total = pick_num(cur_data(), c("bat12_total", "bat_12_total", "burnout")),
    financial_distress = pick_num(cur_data(), c("f10_1", "financial_distress")),
    
    industry_number = pick_num(cur_data(), c("a6", "industry")),
    industry_label = if ("a6" %in% names(itw_all)) {
      as.character(as_factor(a6))
    } else {
      as.character(industry_number)
    },
    
    industry_focus = case_when(
      str_detect(industry_label, regex("retail", ignore_case = TRUE)) | industry_number == 7 ~ "Retail",
      str_detect(industry_label, regex("transport", ignore_case = TRUE)) | industry_number == 9 ~ "Transport",
      TRUE ~ NA_character_
    ),
    
    moderate_severe_k6 = if_else(k6_score >= 13, 1, 0, missing = NA_real_)
  )

# ------------------------------------------------------------
# 5. Domain and hazard scores
# ------------------------------------------------------------

itw_all <- itw_all |>
  mutate(
    Connectedness = score_mean_100(cur_data(), paste0("b1_", 1:10)),
    Leadership = score_mean_100(cur_data(), paste0("b2_", 1:10)),
    Safety = score_mean_100(cur_data(), c(
      "b3x1_1",
      paste0("b3x2_", 1:3),
      paste0("b3x3_", 1:3),
      paste0("b3x4_", 1:3)
    )),
    Work_Design = score_mean_100(cur_data(), c(
      paste0("b4x1_", 1:3),
      paste0("b4x2_", 1:4),
      paste0("b4x3_", 1:3)
    )),
    Capability = score_mean_100(cur_data(), paste0("b5_", 1:7)),
    
    Management_Support = score_mean_100(cur_data(), paste0("b2_", 1:5)),
    Recognition = score_mean_100(cur_data(), paste0("b2_", 6:10)),
    Workload = score_mean_100(cur_data(), paste0("b4x1_", 1:3))
  )

# ------------------------------------------------------------
# 6. Graph summaries
# ------------------------------------------------------------

k6_national <- ci_summary(
  itw_all,
  variable = "k6_score",
  group_vars = c("year", "quarter")
)

burnout_national <- ci_summary(
  itw_all,
  variable = "bat12_total",
  group_vars = c("year", "quarter")
)

industry_k6 <- itw_all |>
  filter(!is.na(industry_focus)) |>
  ci_summary(
    variable = "k6_score",
    group_vars = c("industry_focus", "year", "quarter")
  )

domain_summary <- itw_all |>
  select(year, quarter, Connectedness, Safety, Leadership, Work_Design, Capability) |>
  pivot_longer(
    cols = c(Connectedness, Safety, Leadership, Work_Design, Capability),
    names_to = "domain",
    values_to = "score"
  ) |>
  mutate(domain = str_replace_all(domain, "_", " ")) |>
  group_by(year, domain) |>
  summarise(mean_score = mean(score, na.rm = TRUE), .groups = "drop")

hazard_summary <- itw_all |>
  select(year, quarter, Management_Support, Recognition, Workload) |>
  pivot_longer(
    cols = c(Management_Support, Recognition, Workload),
    names_to = "hazard",
    values_to = "score"
  ) |>
  mutate(hazard = str_replace_all(hazard, "_", " ")) |>
  group_by(year, quarter, hazard) |>
  summarise(mean_score = mean(score, na.rm = TRUE), .groups = "drop")

financial_summary <- itw_all |>
  group_by(year, quarter) |>
  summarise(
    mean_financial_distress = mean(financial_distress, na.rm = TRUE),
    mean_k6 = mean(k6_score, na.rm = TRUE),
    .groups = "drop"
  )

year_cols <- c("2024" = "#00A7C8", "2025" = "#C31861")
```

```{r phase2-prep, include=FALSE}
# ------------------------------------------------------------
# Phase 2 prep: financial distress, age, worker type, utilities, ABS LCI, RBA
# ------------------------------------------------------------

quarter_order <- c(
  "2024 Q1", "2024 Q2", "2024 Q3", "2024 Q4",
  "2025 Q1", "2025 Q2", "2025 Q3", "2025 Q4"
)

itw_phase2 <- itw_all |>
  mutate(
    quarter_full = factor(
      paste(year, quarter),
      levels = quarter_order,
      ordered = TRUE
    ),
    quarter_label = str_replace(as.character(quarter_full), " ", "\n"),
    age_group = if ("age" %in% names(itw_all)) {
      as.character(as_factor(age))
    } else {
      NA_character_
    },
    worker_code = if ("a18" %in% names(itw_all)) {
      as.numeric(a18)
    } else {
      NA_real_
    },
    worker_type = case_when(
      worker_code == 2 ~ "Blue collar",
      worker_code == 1 ~ "White collar",
      TRUE ~ NA_character_
    )
  )

phase2_financial <- itw_phase2 |>
  group_by(quarter_full, quarter_label, year, quarter) |>
  summarise(
    mean_f10 = mean(financial_distress, na.rm = TRUE),
    mean_k6 = mean(k6_score, na.rm = TRUE),
    high_financial_pct = mean(financial_distress >= 8, na.rm = TRUE) * 100,
    n = sum(!is.na(financial_distress)),
    .groups = "drop"
  )

age_financial <- itw_phase2 |>
  filter(!is.na(age_group), !is.na(financial_distress)) |>
  group_by(quarter_full, quarter_label, age_group) |>
  summarise(
    mean_f10 = mean(financial_distress, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

worker_financial <- itw_phase2 |>
  filter(!is.na(worker_type), !is.na(financial_distress)) |>
  group_by(quarter_full, quarter_label, worker_type) |>
  summarise(
    mean_f10 = mean(financial_distress, na.rm = TRUE),
    mean_k6 = mean(k6_score, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

cost_pressure <- itw_phase2 |>
  select(year, quarter, quarter_full, quarter_label, f8_1, f8_2, f8_3) |>
  pivot_longer(
    cols = c(f8_1, f8_2, f8_3),
    names_to = "cost_type",
    values_to = "rating"
  ) |>
  filter(!is.na(rating)) |>
  mutate(
    cost_type = case_when(
      cost_type == "f8_1" ~ "Housing",
      cost_type == "f8_2" ~ "Utilities",
      cost_type == "f8_3" ~ "Food"
    ),
    rating_num = as.numeric(rating),
    difficult = rating_num <= 2
  ) |>
  group_by(quarter_full, quarter_label, cost_type) |>
  summarise(
    difficulty_pct = mean(difficult, na.rm = TRUE) * 100,
    n = n(),
    .groups = "drop"
  )

# ABS Living Cost Index: Employee households
abs_file <- list.files(
  path = ".",
  pattern = "allgroupsbyhouseholds\\.xlsx$",
  recursive = TRUE,
  full.names = TRUE,
  ignore.case = TRUE
)[1]

if (is.na(abs_file)) {
  stop("Could not find allgroupsbyhouseholds.xlsx. Put it in your project folder or data folder.")
}

abs_raw <- read_excel(
  abs_file,
  sheet = "Data1",
  col_names = FALSE
) |>
  clean_names()

abs_lci_2425 <- abs_raw |>
  slice(11:n()) |>
  transmute(
    date_raw = suppressWarnings(as.numeric(x1)),
    employee_lci = suppressWarnings(as.numeric(x3))
  ) |>
  filter(!is.na(date_raw), !is.na(employee_lci)) |>
  mutate(
    date = as.Date(date_raw, origin = "1899-12-30"),
    quarter_full = factor(
      paste(format(date, "%Y"), paste0("Q", quarter(date))),
      levels = quarter_order,
      ordered = TRUE
    ),
    quarter_label = str_replace(as.character(quarter_full), " ", "\n")
  ) |>
  filter(!is.na(quarter_full)) |>
  group_by(quarter_full, quarter_label) |>
  summarise(
    employee_lci = mean(employee_lci, na.rm = TRUE),
    .groups = "drop"
  )

phase2_lci <- phase2_financial |>
  left_join(abs_lci_2425, by = c("quarter_full", "quarter_label"))
  
# RBA cash rate data
rba_rates <- tibble(
  quarter_full = factor(
    c("2024 Q1", "2024 Q2", "2024 Q3", "2024 Q4",
      "2025 Q1", "2025 Q2", "2025 Q3", "2025 Q4"),
    levels = quarter_order,
    ordered = TRUE
  ),
  cash_rate = c(4.35, 4.35, 4.35, 4.35, 4.10, 3.85, 3.60, 3.60)
) |>
  mutate(
    quarter_label = str_replace(as.character(quarter_full), " ", "\n")
  )

phase2_rba <- phase2_financial |>
  left_join(rba_rates, by = c("quarter_full", "quarter_label"))

rba_cor <- cor(
  phase2_rba$mean_f10,
  phase2_rba$cash_rate,
  use = "complete.obs"
)
```





# About SuperFriend

<div class="insight-row">

<div class="insight-card">
<div class="kicker">Organisation</div>
<div class="value">Not-for-profit</div>
<div class="text">SuperFriend works to improve workplace mental health across Australia.</div>
</div>

<div class="insight-card magenta">
<div class="kicker">Mission</div>
<div class="value">Mentally healthy workplaces</div>
<div class="text">The organisation helps workplaces move from awareness to practical action.</div>
</div>

<div class="insight-card orange">
<div class="kicker">Research base</div>
<div class="value">ITW survey</div>
<div class="text">The Indicators of a Thriving Workplace survey tracks workplace mental health and wellbeing nationally.</div>
</div>

</div>

<div class="takeaway">
<strong>In simple terms:</strong> SuperFriend uses workplace mental health research to help organisations understand risk, improve workplace conditions and support workers before problems become severe.
</div>

# Problem and project aim

<div class="insight-row">

<div class="insight-card">
<div class="kicker">Problem</div>
<div class="value">Meaningful change</div>
<div class="text">SuperFriend needed to measure whether changes in ITW variables over time were real and meaningful.</div>
</div>

<div class="insight-card magenta">
<div class="kicker">Why I was required</div>
<div class="value">External expertise</div>
<div class="text">The project needed an intern to help develop the methodology and build internal capability.</div>
</div>

<div class="insight-card orange">
<div class="kicker">Aim</div>
<div class="value">Repeatable method</div>
<div class="text">Create an analysis approach SuperFriend could apply independently to future ITW survey waves.</div>
</div>

</div>

<div class="takeaway">
<strong>Project focus:</strong> Develop a repeatable time-series style methodology to track psychological distress, burnout, workplace domains and psychosocial hazards across quarterly ITW waves in 2024 and 2025.
</div>

<div class="chart-note">
<strong>Important limitation:</strong> The data is cross-sectional, not longitudinal. This means the analysis tracks population-level change across seasonal waves, but does not follow the same individual workers over time.
</div>

# My role and methods

<div class="role-banner">
<div class="role-title">From raw survey data to decision-ready insights</div>
<div class="role-text">
My role was to build a repeatable analysis workflow that could help SuperFriend track workforce mental health changes across repeated ITW survey waves.
</div>
</div>

<div class="method-grid">

<div class="method-step">
<div class="stepno">Step 1</div>
<div class="method-title">Clean + combine</div>
<div class="method-text">Prepared the 2024 and 2025 ITW survey files and created consistent quarterly variables.</div>
</div>

<div class="method-step magenta">
<div class="stepno">Step 2</div>
<div class="method-title">Track change</div>
<div class="method-text">Analysed distress, burnout, workplace domains and psychosocial hazards across quarters.</div>
</div>

<div class="method-step orange">
<div class="stepno">Step 3</div>
<div class="method-title">Test evidence</div>
<div class="method-text">Used quarter comparisons, regression and chi-square tests to avoid overclaiming visual patterns.</div>
</div>

<div class="method-step teal">
<div class="stepno">Step 4</div>
<div class="method-title">Compare industries</div>
<div class="method-text">Focused on Retail and Transport to see whether risk was sector-specific.</div>
</div>

<div class="method-step">
<div class="stepno">Step 5</div>
<div class="method-title">Visualise insights</div>
<div class="method-text">Used violin, bubble, scatter, ranked bars and heatmap charts to communicate findings clearly.</div>
</div>

</div>

<div class="chart-note">
<strong>Method note:</strong> Because the data is repeated cross-sectional rather than longitudinal, the analysis tracks population-level change across waves, not the same workers over time.
</div>

# Statistical tests used 

<div class="test-grid">

<div class="test-card">
<div class="kicker">Quarter differences</div>
<div class="test-title">Pairwise t-tests</div>
<div class="test-text">Compared quarters to identify where distress, burnout, domains and hazards were meaningfully different.</div>
<div class="test-why">Why: shows which quarters changed, not just whether a trend exists.</div>
</div>

<div class="test-card magenta">
<div class="kicker">Multiple testing</div>
<div class="test-title">Bonferroni correction</div>
<div class="test-text">Adjusted p-values because many quarter comparisons were tested.</div>
<div class="test-why">Why: reduces the chance of false significant results.</div>
</div>

<div class="test-card orange">
<div class="kicker">Trend over time</div>
<div class="test-title">Linear regression</div>
<div class="test-text">Tested whether outcomes increased or decreased across quarterly waves.</div>
<div class="test-why">Why: confirms whether visual trends are statistically supported.</div>
</div>

<div class="test-card teal">
<div class="kicker">High-risk group</div>
<div class="test-title">Chi-square + trend test</div>
<div class="test-text">Tested whether the proportion of workers with K6 ≥ 13 changed across quarters.</div>
<div class="test-why">Why: focuses on clinically meaningful distress risk.</div>
</div>

<div class="test-card">
<div class="kicker">Economic links</div>
<div class="test-title">Correlation tests</div>
<div class="test-text">Checked relationships between financial distress, psychological distress, ABS LCI and RBA rates.</div>
<div class="test-why">Why: connects worker wellbeing with economic pressure.</div>
</div>

<div class="test-card magenta">
<div class="kicker">Group differences</div>
<div class="test-title">Interaction regression</div>
<div class="test-text">Compared whether blue-collar and white-collar workers responded differently to interest-rate pressure.</div>
<div class="test-why">Why: shows whether financial pressure affects worker groups differently.</div>
</div>

</div>

<div class="takeaway">
<strong>Overall purpose:</strong> These tests helped separate meaningful workforce signals from normal quarterly variation and explain why some groups may need more targeted support.
</div>


# Psychological distress peaks around Q2

<div class="two-col">

<div class="chart-frame">

```{r k6-trend-line-final, fig.width=8.6, fig.height=4.8}
k6_trend <- k6_national |>
  mutate(year = factor(year))

p_trend <- ggplot(
  k6_trend,
  aes(x = quarter, y = mean, group = year, colour = year, fill = year)
) +
  annotate(
    "rect",
    xmin = 1.65,
    xmax = 2.35,
    ymin = -Inf,
    ymax = Inf,
    fill = "#F49E22",
    alpha = .14
  ) +
  geom_ribbon(
    aes(ymin = lower, ymax = upper),
    alpha = .16,
    colour = NA
  ) +
  geom_line(linewidth = 2.2, lineend = "round") +
  geom_point(size = 5.5) +
  geom_point(
    data = k6_trend |> filter(quarter == "Q2"),
    shape = 21,
    size = 9,
    stroke = 2.3,
    fill = "white"
  ) +
  annotate(
    "label",
    x = 2,
    y = max(k6_trend$upper, na.rm = TRUE) + .18,
    label = "Q2 pressure point",
    fill = "#173A59",
    colour = "white",
    fontface = "bold",
    size = 3.8
  ) +
  scale_colour_manual(values = year_cols) +
  scale_fill_manual(values = year_cols) +
  labs(
    subtitle = "Mean Psychological distress score by quarter with 95% confidence interval",
    x = NULL,
    y = "Mean psychological distress score"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.title.y = element_text(face = "bold", colour = "#173A59"),
    axis.text = element_text(colour = "#16212F"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank(),
    plot.margin = margin(6, 10, 2, 10)
  )

p_trend
```

</div>

<div class="side-card">
<div class="kicker">Main trend</div>
<div class="number">Q2 peak</div>
<div class="label">Psychological distress rises around the middle of the year, suggesting a predictable pressure point for prevention.</div>
</div>

</div>

# Distress distribution by quarter

<div class="insight-row">

<div class="insight-card">
<div class="kicker">View</div>
<div class="value">Distribution</div>
<div class="text">Shows the full spread of Psychological distress scores, not only the average.</div>
</div>

<div class="insight-card magenta">
<div class="kicker">Risk line</div>
<div class="value">Psychological distress ≥ 13</div>
<div class="text">Marks the moderate-to-severe psychological distress threshold.</div>
</div>

<div class="insight-card orange">
<div class="kicker">Message</div>
<div class="value">High-risk tail</div>
<div class="text">The upper tail highlights workers reporting much higher distress.</div>
</div>

</div>

<div class="chart-frame">

```{r k6-violin-fancy, fig.width=11, fig.height=4.2}
violin_data <- itw_all |>
  filter(!is.na(k6_score), !is.na(quarter), !is.na(year)) |>
  mutate(year = factor(year))

p1 <- ggplot(
  violin_data,
  aes(x = quarter, y = k6_score, fill = year)
) +
  geom_violin(
    position = position_dodge(width = .82),
    alpha = .55,
    trim = TRUE,
    colour = NA
  ) +
  geom_boxplot(
    position = position_dodge(width = .82),
    width = .13,
    outlier.alpha = .06,
    alpha = .88,
    colour = "#173A59",
    linewidth = .55
  ) +
  geom_hline(
    yintercept = 13,
    linetype = "dashed",
    linewidth = 1.1,
    colour = "#C31861"
  ) +
  annotate(
    "label",
    x = 3.55,
    y = 14.4,
    label = "K6 ≥ 13 threshold",
    fill = "#C31861",
    colour = "white",
    fontface = "bold",
    size = 3.8
  ) +
  scale_fill_manual(values = year_cols) +
  labs(
    subtitle = "Violin width shows where workers are concentrated",
    x = "Quarter",
    y = "Psychological distress score"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.title.y = element_text(face = "bold", colour = "#173A59"),
    axis.text = element_text(colour = "#16212F"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank(),
    plot.margin = margin(6, 10, 2, 10)
  )

p1
```

</div>

# Distress and burnout tell different stories

<div class="two-col">

<div class="chart-frame">

```{r k6-burnout-bubble-fancy, fig.width=8.6, fig.height=4.8}
bubble_data <- itw_all |>
  group_by(year, quarter) |>
  summarise(
    mean_k6 = mean(k6_score, na.rm = TRUE),
    mean_burnout = mean(bat12_total, na.rm = TRUE),
    n = sum(!is.na(k6_score) & !is.na(bat12_total)),
    .groups = "drop"
  ) |>
  mutate(
    year = factor(year),
    quarter_label = as.character(quarter)
  )

k6_mid <- mean(bubble_data$mean_k6, na.rm = TRUE)
burnout_mid <- mean(bubble_data$mean_burnout, na.rm = TRUE)

p2 <- ggplot(
  bubble_data,
  aes(
    x = mean_burnout,
    y = mean_k6,
    size = n,
    colour = year
  )
) +
  annotate(
    "rect",
    xmin = burnout_mid,
    xmax = Inf,
    ymin = k6_mid,
    ymax = Inf,
    fill = "#C31861",
    alpha = .07
  ) +
  annotate(
    "rect",
    xmin = -Inf,
    xmax = burnout_mid,
    ymin = -Inf,
    ymax = k6_mid,
    fill = "#00A7C8",
    alpha = .06
  ) +
  geom_vline(
    xintercept = burnout_mid,
    linetype = "dashed",
    linewidth = .9,
    colour = "#64748B"
  ) +
  geom_hline(
    yintercept = k6_mid,
    linetype = "dashed",
    linewidth = .9,
    colour = "#64748B"
  ) +
  geom_point(alpha = .82) +
  geom_text(
    aes(label = quarter_label),
    colour = "white",
    fontface = "bold",
    size = 4.2
  ) +
  scale_colour_manual(values = year_cols) +
  scale_size(range = c(12, 22), guide = "none") +
  labs(
    subtitle = "Each bubble is one quarter; size reflects sample size",
    x = "Mean burnout",
    y = "Mean psychological distress"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.title = element_text(face = "bold", colour = "#173A59"),
    axis.text = element_text(colour = "#16212F"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank(),
    plot.margin = margin(6, 10, 2, 10)
  )

p2
```

</div>

<div class="side-card">
<div class="kicker">Interpretation</div>
<div class="number">2 signals</div>
<div class="label">Psychological distress captures sharper distress, while burnout reflects slower accumulated strain.</div>
</div>

</div>

# Retail and Transport have different risk profiles

<div class="two-col">

<div class="chart-frame">

```{r industry-profile-scatter-fancy, fig.width=8.6, fig.height=4.8}
industry_profile <- itw_all |>
  filter(!is.na(industry_focus)) |>
  group_by(industry_focus, year, quarter) |>
  summarise(
    mean_k6 = mean(k6_score, na.rm = TRUE),
    mean_burnout = mean(bat12_total, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  mutate(year = factor(year))

p4 <- ggplot(
  industry_profile,
  aes(
    x = mean_burnout,
    y = mean_k6,
    colour = industry_focus,
    shape = year
  )
) +
  annotate(
    "rect",
    xmin = median(industry_profile$mean_burnout, na.rm = TRUE),
    xmax = Inf,
    ymin = median(industry_profile$mean_k6, na.rm = TRUE),
    ymax = Inf,
    fill = "#C31861",
    alpha = .07
  ) +
  geom_point(size = 6, alpha = .9) +
  geom_text(
    aes(label = quarter),
    vjust = -1,
    fontface = "bold",
    colour = "#173A59",
    size = 3.7
  ) +
  scale_colour_manual(values = c("Retail" = "#C31861", "Transport" = "#00A7C8")) +
  labs(
    subtitle = "Each point is an industry-quarter average",
    x = "Mean burnout",
    y = "Mean psychological distress"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.title = element_text(face = "bold", colour = "#173A59"),
    axis.text = element_text(colour = "#16212F"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank(),
    plot.margin = margin(6, 10, 2, 10)
  )

p4
```

</div>

<div class="side-card">
<div class="kicker">Industry view</div>
<div class="number">Retail risk</div>
<div class="label">Retail tends to sit higher on distress, supporting industry-specific workplace action.</div>
</div>

</div>

# Workplace domains show strengths and gaps

<div class="two-col">

<div class="chart-frame">

```{r domain-strength-gap-final, fig.width=8.6, fig.height=4.8}
domain_rank <- domain_summary |>
  mutate(year = as.character(year)) |>
  select(year, domain, mean_score) |>
  pivot_wider(names_from = year, values_from = mean_score) |>
  mutate(
    change = `2025` - `2024`,
    change_label = paste0(if_else(change >= 0, "+", ""), round(change, 1)),
    domain = fct_reorder(domain, `2025`)
  )

p5 <- ggplot(domain_rank, aes(y = domain)) +
  geom_segment(
    aes(x = 55, xend = `2025`, yend = domain),
    linewidth = 13,
    colour = "#C31861",
    alpha = .18,
    lineend = "round"
  ) +
  geom_segment(
    aes(x = `2024`, xend = `2025`, yend = domain),
    linewidth = 3.2,
    colour = "#173A59",
    alpha = .35,
    lineend = "round"
  ) +
  geom_point(
    aes(x = `2024`),
    size = 5.8,
    colour = "#00A7C8",
    alpha = .95
  ) +
  geom_point(
    aes(x = `2025`),
    size = 7.2,
    colour = "#C31861",
    alpha = .95
  ) +
  geom_text(
    aes(x = `2025` + .7, label = round(`2025`, 1)),
    colour = "#173A59",
    fontface = "bold",
    size = 4.2,
    hjust = 0
  ) +
  geom_label(
    aes(x = 80.5, label = change_label),
    fill = "white",
    colour = "#173A59",
    fontface = "bold",
    size = 3.7,
    label.size = 0
  ) +
  annotate(
    "text",
    x = 56.4,
    y = length(unique(domain_rank$domain)) + .55,
    label = "2024",
    colour = "#00A7C8",
    fontface = "bold",
    size = 4.2
  ) +
  annotate(
    "text",
    x = 60.2,
    y = length(unique(domain_rank$domain)) + .55,
    label = "2025",
    colour = "#C31861",
    fontface = "bold",
    size = 4.2
  ) +
  annotate(
    "text",
    x = 80.5,
    y = length(unique(domain_rank$domain)) + .55,
    label = "change",
    colour = "#173A59",
    fontface = "bold",
    size = 4.2
  ) +
  labs(
    subtitle = "0–100 scale: higher scores indicate stronger workplace performance / better management",
    x = "Mean domain score / 100",
    y = NULL
  ) +
  coord_cartesian(xlim = c(55, 83), clip = "off") +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.title.x = element_text(face = "bold", colour = "#173A59"),
    axis.text.y = element_text(face = "bold", colour = "#173A59"),
    axis.text.x = element_text(colour = "#16212F"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(6, 45, 2, 10)
  )

p5
```

</div>

<div class="side-card">
<div class="kicker">Workplace levers</div>
<div class="number">Strengths + Gaps</div>
<div class="label">Connectedness and Safety are stronger areas, while Capability, Work Design and Leadership show where support can be improved.</div>
</div>

</div>

# Psychosocial hazards highlight workplace pressure

<div class="two-col">

<div class="chart-frame">

```{r hazard-clean-heatmap, fig.width=8.6, fig.height=4.8}
hazard_tile_data <- hazard_summary |>
  mutate(
    year = factor(year),
    hazard = fct_relevel(hazard, "Management Support", "Recognition", "Workload"),
    label = round(mean_score, 1)
  )

p6 <- ggplot(
  hazard_tile_data,
  aes(x = quarter, y = hazard, fill = mean_score)
) +
  geom_tile(colour = "white", linewidth = 1.5) +
  geom_text(
    aes(label = label),
    colour = "white",
    fontface = "bold",
    size = 4.5
  ) +
  facet_wrap(~ year, nrow = 1) +
  scale_fill_gradientn(
    colours = c("#C31861", "#F49E22", "#17B8A6", "#00A7C8"),
    name = "Score"
  ) +
  labs(
    subtitle = "0–100 scale: higher scores indicate better management of psychosocial hazards",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.text.y = element_text(face = "bold", colour = "#173A59"),
    axis.text.x = element_text(colour = "#16212F"),
    strip.text = element_text(face = "bold", colour = "#173A59", size = 14),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(face = "bold", colour = "#173A59"),
    plot.margin = margin(6, 10, 2, 10)
  )

p6
```

</div>

<div class="side-card">
<div class="kicker">Main message</div>
<div class="number">Not just Workload</div>
<div class="label">Support, recognition and leadership quality are central workplace levers.</div>
</div>

</div>

# Financial distress and economic pressure

<div class="phase-band">
<p>Exploring how financial distress connects with psychological distress, living costs, age, occupation, essential costs and RBA interest rates.</p>
</div>

<div class="insight-row" style="margin-top:24px;">

<div class="insight-card">
<div class="kicker">Focus</div>
<div class="value">Financial distress</div>
<div class="text">How workers experience cost-of-living and money pressure.</div>
</div>

<div class="insight-card magenta">
<div class="kicker">Groups</div>
<div class="value">Age + occupation</div>
<div class="text">Compare financial pressure across age and blue/white collar workers.</div>
</div>

<div class="insight-card orange">
<div class="kicker">Context</div>
<div class="value">LCI + RBA</div>
<div class="text">Connect worker distress with living costs and interest rates.</div>
</div>

</div>

# Financial pressure links to distress

<div class="two-col">

<div class="chart-frame">

```{r financial-k6-scatter-fancy, fig.width=8.6, fig.height=4.8}
scatter_data <- itw_all |>
  filter(!is.na(financial_distress), !is.na(k6_score), !is.na(year)) |>
  sample_n(size = min(3500, n())) |>
  mutate(year = factor(year))

mean_by_finance <- itw_all |>
  filter(!is.na(financial_distress), !is.na(k6_score), !is.na(year)) |>
  group_by(year, financial_distress) |>
  summarise(
    mean_k6 = mean(k6_score, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  mutate(year = factor(year))

p3 <- ggplot() +
  geom_jitter(
    data = scatter_data,
    aes(
      x = financial_distress,
      y = k6_score,
      colour = year
    ),
    alpha = .10,
    size = 1.8,
    width = .16,
    height = .18
  ) +
  geom_smooth(
    data = scatter_data,
    aes(
      x = financial_distress,
      y = k6_score,
      colour = year
    ),
    method = "lm",
    se = TRUE,
    linewidth = 1.6,
    alpha = .18
  ) +
  geom_point(
    data = mean_by_finance,
    aes(
      x = financial_distress,
      y = mean_k6,
      colour = year,
      size = n
    ),
    alpha = .95
  ) +
  geom_line(
    data = mean_by_finance,
    aes(
      x = financial_distress,
      y = mean_k6,
      colour = year,
      group = year
    ),
    linewidth = 1.4
  ) +
  geom_hline(
    yintercept = 13,
    linetype = "dashed",
    linewidth = 1,
    colour = "#C31861"
  ) +
  annotate(
    "label",
    x = max(scatter_data$financial_distress, na.rm = TRUE) * .72,
    y = 14.3,
    label = "K6 ≥ 13 threshold",
    fill = "#C31861",
    colour = "white",
    fontface = "bold",
    size = 3.6
  ) +
  scale_colour_manual(values = year_cols) +
  scale_size(range = c(4, 11), guide = "none") +
  labs(
    subtitle = "Faint dots are workers, large points show average K6 at each financial distress level",
    x = "Financial distress score",
    y = "Psychological distress score"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.title = element_text(face = "bold", colour = "#173A59"),
    axis.text = element_text(colour = "#16212F"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank(),
    plot.margin = margin(6, 10, 2, 10)
  )

p3
```

</div>

<div class="side-card">
<div class="kicker">Phase 2 link</div>
<div class="number">Positive Trend</div>
<div class="label">Higher financial distress is associated with higher psychological distress.</div>
</div>

</div>

# Living costs provide economic context

<div class="two-col">

<div class="chart-frame">

```{r abs-lci-financial-distress, fig.width=8.6, fig.height=4.8}
phase2_lci_plot <- phase2_lci |>
  mutate(
    lci_scaled = employee_lci / 20
  )

p_lci <- ggplot(phase2_lci_plot, aes(x = quarter_full)) +
  geom_line(
    aes(y = mean_f10, group = 1, colour = "Financial distress"),
    linewidth = 1.8,
    lineend = "round"
  ) +
  geom_point(
    aes(y = mean_f10, colour = "Financial distress"),
    size = 4
  ) +
  geom_line(
    aes(y = lci_scaled, group = 1, colour = "ABS Living Cost Index (scaled ÷20)"),
    linewidth = 1.8,
    lineend = "round"
  ) +
  geom_point(
    aes(y = lci_scaled, colour = "ABS Living Cost Index (scaled ÷20)"),
    size = 4
  ) +
  scale_colour_manual(
    values = c(
      "Financial distress" = "#1E40FF",
      "ABS Living Cost Index (scaled ÷20)" = "#E31A1C"
    )
  ) +
  labs(
    subtitle = "Financial distress compared with ABS Employee households LCI",
    x = "Quarter",
    y = "Financial distress / scaled LCI",
    colour = NULL
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.title.y = element_text(face = "bold", colour = "#173A59"),
    axis.title.x = element_text(face = "bold", colour = "#173A59"),
    axis.text.x = element_text(face = "bold", colour = "#173A59"),
    axis.text.y = element_text(colour = "#16212F"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank(),
    plot.margin = margin(6, 10, 2, 10)
  )

p_lci
```

</div>

<div class="side-card">
<div class="kicker">Cost-of-living context</div>
<div class="number">ABS LCI</div>
<div class="label">The Employee households Living Cost Index helps interpret financial distress against broader living-cost pressure.</div>
</div>

</div>

# Financial distress differs by age group

<div class="two-col">

<div class="chart-frame">

```{r age-financial-heatmap, fig.width=8.6, fig.height=4.8}
p_age <- ggplot(
  age_financial,
  aes(
    x = quarter_full,
    y = fct_rev(age_group),
    fill = mean_f10
  )
) +
  geom_tile(colour = "white", linewidth = 1.2) +
  geom_text(
    aes(label = round(mean_f10, 1)),
    colour = "white",
    fontface = "bold",
    size = 3.6
  ) +
  scale_fill_gradientn(
    colours = c("#00A7C8", "#17B8A6", "#F49E22", "#C31861"),
    name = "Mean Financial distress"
  ) +
  labs(
    subtitle = "Mean financial distress by age group and quarter",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.text.x = element_text(face = "bold", colour = "#173A59"),
    axis.text.y = element_text(face = "bold", colour = "#173A59"),
    panel.grid = element_blank(),
    legend.position = "bottom",
    plot.margin = margin(6, 10, 2, 10)
  )

p_age
```

</div>

<div class="side-card">
<div class="kicker">Age profile</div>
<div class="number">Younger Workers</div>
<div class="label">Age categories show whether financial pressure is concentrated among younger or working-age groups.</div>
</div>

</div>

# Financial pressure differs by occupation type

<div class="two-col">

<div class="chart-frame">

```{r worker-financial-line, fig.width=8.6, fig.height=4.8}
p_worker <- ggplot(
  worker_financial,
  aes(
    x = quarter_full,
    y = mean_f10,
    colour = worker_type,
    group = worker_type
  )
) +
  geom_line(linewidth = 2.1, lineend = "round") +
  geom_point(size = 5) +
  geom_text(
    aes(label = round(mean_f10, 1)),
    vjust = -1,
    fontface = "bold",
    colour = "#173A59",
    size = 3.5
  ) +
  scale_colour_manual(
    values = c("Blue collar" = "#C31861", "White collar" = "#00A7C8")
  ) +
  labs(
    subtitle = "Mean financial distress by occupational group",
    x = NULL,
    y = "Mean financial distress score"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.title.y = element_text(face = "bold", colour = "#173A59"),
    axis.text.x = element_text(face = "bold", colour = "#173A59"),
    axis.text.y = element_text(colour = "#16212F"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.title = element_blank(),
    plot.margin = margin(6, 10, 2, 10)
  )

p_worker
```

</div>

<div class="side-card">
<div class="kicker">Occupation</div>
<div class="number">Blue vs White</div>
<div class="label">This shows whether financial pressure differs by type of work, not just across the whole workforce.</div>
</div>

</div>


# RBA interest rates provide economic context

<div class="two-col">

<div class="chart-frame">

```{r rba-financial-distress, fig.width=8.6, fig.height=4.8}
phase2_rba_long <- phase2_rba |>
  select(quarter_full, mean_f10, cash_rate) |>
  pivot_longer(
    cols = c(mean_f10, cash_rate),
    names_to = "indicator",
    values_to = "value"
  ) |>
  mutate(
    indicator = case_when(
      indicator == "mean_f10" ~ "Financial Distress",
      indicator == "cash_rate" ~ "RBA Cash Rate (%)"
    )
  )

p_rba <- ggplot() +
  geom_line(
    data = phase2_rba_long |> filter(indicator == "Financial Distress"),
    aes(x = quarter_full, y = value, group = indicator, colour = indicator),
    linewidth = 1.8
  ) +
  geom_point(
    data = phase2_rba_long |> filter(indicator == "Financial Distress"),
    aes(x = quarter_full, y = value, colour = indicator),
    size = 4
  ) +
  geom_line(
    data = phase2_rba_long |> filter(indicator == "RBA Cash Rate (%)"),
    aes(x = quarter_full, y = value, group = indicator, colour = indicator),
    linewidth = 1.5,
    linetype = "dashed"
  ) +
  geom_point(
    data = phase2_rba_long |> filter(indicator == "RBA Cash Rate (%)"),
    aes(x = quarter_full, y = value, colour = indicator),
    size = 3.5
  ) +
  scale_colour_manual(
    values = c(
      "Financial Distress" = "#1E40FF",
      "RBA Cash Rate (%)" = "#F49E22"
    )
  ) +
  labs(
    subtitle = "Comparing Workforce Sentiment with Monetary Policy",
    x = "Quarter",
    y = "Score",
    colour = "Indicator"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.background = element_rect(fill = "#F7FAFC", colour = NA),
    panel.background = element_rect(fill = "#F7FAFC", colour = NA),
    plot.subtitle = element_text(colour = "#435469", size = 12),
    axis.title.y = element_text(face = "bold", colour = "#173A59"),
    axis.title.x = element_text(face = "bold", colour = "#173A59"),
    axis.text.x = element_text(face = "bold", colour = "#173A59", angle = 45, hjust = 1),
    axis.text.y = element_text(colour = "#16212F"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(face = "bold"),
    plot.margin = margin(6, 10, 2, 10)
  )

p_rba
```

</div>

<div class="side-card">
<div class="kicker">Macro context</div>
<div class="number">RBA rates</div>
<div class="label">Interest rates provide economic context for interpreting changes in workforce financial distress.</div>
</div>

</div>

# Final takeaways

<div class="chart-note">
<strong>Project objective:</strong> Develop a repeatable approach to identify meaningful workforce mental health and financial distress signals across ITW survey waves, and translate those signals into practical workplace action.
</div>

<div class="insight-row">

<div class="insight-card">
<div class="kicker">When to act</div>
<div class="value">Distress peaks around Q2</div>
<div class="text">This helps SuperFriend identify a recurring mid-year pressure point where early support may be useful.</div>
</div>

<div class="insight-card magenta">
<div class="kicker">Who to support</div>
<div class="value">Risk is uneven</div>
<div class="text">Industry, age and occupation results show that risk is not spread equally across all workers.</div>
</div>

<div class="insight-card orange">
<div class="kicker">What to consider</div>
<div class="value">Economic context matters</div>
<div class="text">Financial distress should be interpreted alongside psychological distress, living costs and interest rates.</div>
</div>

</div>

<div class="takeaway">
<strong>Overall conclusion:</strong> This project developed a repeatable cross-sectional trend approach that helps SuperFriend separate meaningful workforce mental health signals from normal quarterly variation, and translate those signals into targeted workplace action.
</div>